### 2019-04-28 neural_frontend/doc/main.tex
updates gitignore

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
%\usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{tikz,placeins}
\usepackage{subcaption}
%\usepackage[caption = false]{subfig}
\usepackage{amssymb,amsmath,bm}
\usetikzlibrary{arrows,decorations.markings}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\large]

\graphicspath{{../figures/}}

\title{WEVE: Weighted Enhancement Using Variance Estimation}

%\renewcommand{\vec}[1]{\mathbf{#1}}

\newcommand{\chanwcom}{{C. Kim}}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.85}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  David S.~Hippocampus\thanks{Use footnote for providing further
    information about author (webpage, alternative
    address)---\emph{not} for acknowledging funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
This paper presents a new robust neural frontend layer referred to as Power
Normalized Neural Frontend (PNNF) for robust speech recognition. PNNF
is implemented as a layer in an end-to-end all neural-network speech 
recognition system. For speech recognition, features such as Mel Filterbank
Cepstral Coefficient (MFCC) or log-Mel have been used very widely.
More recently, features motivated by human auditory processing have shown
improvement under certain conditions. Nevertheless, these auditory
features are constructed using signal processing blocks with hand-crafted
parameters for certain conditions. Therefore, the performance is not
always optimal for various kind of acoustic conditions.
 PNNF frontend layer is loosely motivated by human auditory system.
This processing uses the concept of the auditory masking and the auditory nonlinearity.
However, instead of directly using theories and parameters about human auditory masking,
neural networks make decision about it from the training data.
Even though the structure of the PNNF layer is very simple, this layer
significantly boosts the performance of the speech recognition system.
\end{abstract}

\section{Introduction}

\section{Feature enhancement}

In the feature enhancement process, we estimate a clean feature 
$\vec{y}$ given a corrupt feature $\vec{x}$.

In this case, the following $L_2$ norm is widely used:
\begin{align}
  l(\vec{y}, \vec{x}) = 
\end{align}

Feature mapping is a process of mapping noisy feature vectors
to clean feature vectors to enhance robustness
\cite{C_Kim_ASRU_2009_1, k_han_taslp_2015_00, k_han_icassp_2014_00}.
This procedure may be formulated by the following equation:
\begin{align}
  Y() = X()
\end{align}

From noisy features, we may infer 
Another

Masking and feature mapping are closely related techniques.
In feature mapping, enhanced features are predicted from
noisy features using machine learning models
The masking-based approaches with
Joint Adaptive Training (JAT) have been shown
to be effective to a large training set
\cite{a_narayanan_interspeech_2015_1}.

\section{Structure of the neural front-end}

\subsection{Neural nonlinearity}
In this section, we describe our neural nonlinearity algorithm. 
\begin{align}
  \vec{q[m, l]} = \vec{x[m, l]} ^ {\vec{\alpha}} 
\end{align}

\subsection{Neural background removal}
In this section, we describe our Neural Background Removal (NBR) approach.
Fig. \label{fig:nbr_diagram} shows the structure of the NBR processing.

$\vec{q}$ is the non-linearity output defined in

\begin{figure}
    \centering
    \resizebox{80mm}{!}{%
         \input{../figures/background_removal.tex}} %
    \caption {
      The structure of the Neural Background Removal (NBR) block.
      \label{fig:nbr_diagam}
    }
    %\end{center}
\end{figure}

\subsection{  }

NIPS requires electronic submissions.  The electronic submission site
is
\begin{center}
  \url{https://cmt.research.microsoft.com/NIPS2017/}
\end{center}

Please read carefully the instructions below and follow them
faithfully.

\section{The structure of the SPL algorithm}

Fig. \label{fig:entire_diagram} shows the block diagram for
enhancing the noisy feature.

%\begin{figure}
%    \centering
%    \resizebox{80mm}{!}{%
%         \input{../figures/entire_diagram.tex}} %
%    \caption {
%      The structure of feature enhancement using peak
%      value normalization and LSTMs.
%      \label{fig:entire_diagam}
%    }
%    %\end{center}
%\end{figure}

Fig. \ref{fig:entire_diagram} shows the structure for enhancing
noisy feature.

\subsection{Simulated utterance generation}
We use the room simulation system introduced in
\cite{C_Kim_INTERSPEECH_2017_1} to generate noisy speech
from clean speech.
\begin{align}
  y[n]
\end{align}
For 
\subsection{Review of the Room Simulator for Data Augmentation}
%
%
%\begin{figure}
%  \centering
%    \begin{subfigure}{\linewidth}
%      \centering
%      \includegraphics[width=70mm]{room_diagram.eps}
%      \caption{}
%      \label{fig:room_diagram}
%    \end{subfigure}
%   \\
%    \begin{subfigure}{\linewidth}
%      \centering
%      \includegraphics[width=70mm]{image_method.eps}
%      \caption{}
%      \label{fig:image_method}
%    \end{subfigure}
%    \caption{
%    (a) A simulated room: There may be multiple microphones, a single target
%    sound source, multiple noise sources in a cuboid-shape room with
%    acoustically reflective walls. (b) A diagram showing the location
%    of the real and the virtual sound sources assuming that the walls
%    reflects acoustic wave like a mirror \cite{C_Kim_INTERSPEECH_2017_1}.}
%  \vspace{-10mm}
%\end{figure}
%
%
In this section, we briefly review the structure of the Google
Room Simulator for generating simulated utterances
to train acoustic models for speech recognition systems
\cite{C_Kim_INTERSPEECH_2017_1}. We assume a room of a
rectangular cuboid-shape as shown in Fig. \ref{fig:room_diagram}.
Assuming that all the walls of a room reflect acoustically uniformly,
we use the image method to model the Room Impulse Responses (RIRs)
\cite{J_Allen_JASA_1979, E_A_Lehmann_ASPAA_2007, S_G_McGovern_RIR}.

Fig. \ref{fig:image_method} shows how virtual sources
appearing as empty circles are constructed in the image method.
Even though the virtual rooms appear to be created in two-dimensions
in Fig. \ref{fig:image_method}, in the ``\emph{Room Simulator}",
they are constructed in three-dimensions. For example, in our
work for training the acoustic model for Google Home, we consider
$17 \times 17 \times 17 = 4913$ virtual rooms for RIR calculation
\cite{C_Kim_INTERSPEECH_2017_1}.
Following the image method, the impulse response is calculated
using the following equation
\cite{ J_Allen_JASA_1979, E_A_Lehmann_ASPAA_2007}:
\begin{align}
    h[n] = \sum_{i = 0}^{I-1} \frac{r^{g_i}}{d_i}
    \delta \left[n -\left \lceil{\frac{d_i f_s}{c_0}}\right \rceil \right],
      \label{eq:h_n_calculation}
\end{align}
where $i$ is the index of each virtual sound source, and $d_i$ is the distance
from that sound source to the microphone, $r$ is
the reflection coefficient of the wall, $g_i$ is
the number of the reflections
to that sound source, $f_s$ is the sampling rate of the RIR, and $c_0$ is the
speed of sound in the air. We use the value of $f_s = 16,000$ \textit{Hz} and
$c_0 = 343$ \textit{m/s} for the room simulator \cite{B_Li_INTERSPEECH_2017_1}.
For $d_i$, $r$, we use numbers created by a random number generator
following specified distributions for each impulse response \cite{C_Kim_INTERSPEECH_2017_1}.

Assuming that there are $I$ sound sources including one target
source and $J$ microphones, the received signal at microphone
$j$ is given by:
\begin{align}
  y_j[n] =  \sum_{i=0}^{I-1} \alpha_{ij} \left(h_{ij}[n] * x_i[n]\right).
  \label{eq:y_j_def}
\end{align}
Since we used a two-microphones system in
\cite{B_Li_INTERSPEECH_2017_1, C_Kim_INTERSPEECH_2017_1},
$J$ is two, and for the number noise sources, we used a value from zero
up to three \cite{C_Kim_INTERSPEECH_2017_1}.
%

\subsection{Feature extraction}

Even though the log-mel coefficients have been widely used
as features for speech recognition \cite{sbdavisandpm1980},
the log non-linearity has a disadvantage in that the value
diverges to negative infinity as the mel filterbank coefficient
approaches zero \cite{C_Kim_IEEETran_2016_1}.
Thus, we use the power-law nonlinearity
rather than the log-nonlinearity:
\begin{align}
  q[m, l] =  (p[m, l])^{\frac{1}{15}}
  \label{eq:power_law_nonlineartiy}
\end{align}
We use the power coefficient of $\frac{1}{15}$ as in
\cite{C_Kim_IEEETran_2016_1, C_Kim_ICASSP_2012_1}.

\subsection{Feature normalization using the peak-value}

The power-mel coefficients $q_x[m, l]$ and $q_y[m, l]$
are obtained from the clean and the simulated noisy
speech $x[n]$ and $y[n]$ respectively. $q_{x, \text{peak}}$
and $q_{y, \text{peak}}$

\section{Estimation of the expected value and the variance
  of the enhanced feature from the noisy feature}
\label{sec:estimation}

\begin{figure}
    \centering
    \resizebox{80mm}{!}{%
         \input{../figures/training.tex}} %
    \caption {
      Training of neural networks to estimate the clean feature and error ratio from corrupt feature
      \label{MFCCPNCCComparison}
    }
    %\end{center}
\end{figure}

In this section, we discuss how to predict the clean feature and the
error ratio from the corrupt feature. Let us denote the clean target feature,
the corrupt feature by $\boldsymbol{x}, \boldsymbol{y}$ respectively.

Using the room simulation system described in \cite{C_Kim_INTERSPEECH_2017_1},
we create a pair of clean and
corrupt utterances. Log-mel [XX] features are calculated from these original clean
and corrupt utterances. Let us denote the clean feature by $\boldsymbol{t}_i$ and
the simulated corrupt feature by $\boldsymbol{x}_i$, respectively where $i$ is the
utterance index. The training set is represented by the following set:
\begin{align}
  \mathcal{T} = \{<\boldsymbol{x}_i, \boldsymbol{t}_i>| 0 \le i \le N - 1\}
\end{align}
where $N$ is the number of training examples.

In the ECE algorithm, we first estimate the true target $\boldsymbol{t}_i$. The
estimation error is defined by the following equation:
\begin{align}
  \boldsymbol{e}_i =  \boldsymbol{t}_i - \boldsymbol{x}_i, \quad 0 \le i \le N - 1
\end{align}

The norm of the error vector $\boldsymbol{e}_i$ would be generally large when
the norm of the estimated clean feature is large.

Thus, instead of trying to estimate the error itself, the neural network tries to
estimate the error ratio given as follows:

\begin{align}
  r_i = {}{}
\end{align}

\section{Generation of simulated clean and noisy data sets}
\label{gen_inst}

To train neural networks described in section \ref{sec:estimation}, we
need a training set consisting of pairs of clean speech and noisy speech.
Unfortunately, it is not easy to have such a training set from real speech
utterances. Thus, we use the simulation system described in
\cite{C_Kim_INTERSPEECH_2017_1} to synthetically generate noisy speech
utterances from the clean speech utterances.

In our application, for clean training set, we used Wall Street Journal (WSJ)
si-284. For noise set, we used the Resource Management 1 (RM1)
\cite{price_p_ldc_1993} 50 \% of time 
and DEMAND noise [TODO(chanwcom) Adds reference] for the remaining 50 \% of time.

TODO(chanwcom) Cite DEMAND noise set.

%
%
%
%\begin{figure*}
%  \centering
%    \begin{subfigure}[\label{fig:randomly_selected_grid_coordinates}]{0.5\textwidth} {
%        \includegraphics[width=\textwidth]{target_estimation}}
%    \caption{Grid from random-sampling}
%    \end{subfigure}
%   % \begin{subfigure}[\label{fig:intersection_point}]{0.5\textwidth} {
%   %   \includegraphics[width=\textwidth]{intersection_point}}
%   % \caption{Calculation of grid coordinates}
%   % \end{subfigure}
%  \caption {Random Grid Sampling}
%\end{figure*}
%
%
%
\section{Estimation of the optimal weight in the EFW algorithm}
In this section, we describe the procedure of feature weighting approach.
Suppose that the feature before enhancement is represented by $\vec{x}$.
Let us represent the inference neural network to estimate the target by
$\mathcal{T}$.

\begin{align}
  \mathcal{X} = \{  \}
\end{align}

Instead of directly using the enhanced feature $\widehat{\vec{y}}$, we consider
the following interpolated feature $\bm{z}$ from the original corrupt
input $\bm{x}$ and .
\begin{align}
  \vec{z} = \vec{w} \odot \widehat{\vec{y}} \ + (1 - \vec{w}) \odot \vec{x} 
    \label{eq:vec_z}
\end{align}
where $\odot$ denotes the Hadamard product (entry-wise product).

Let us denote the estimated variance vector by $\hat{\vec{v}}$:
\begin{align}
  \hat{\vec{v}} & = \text{Var}[\widehat{\vec{y}}] \nonumber \\
                & = E[(y - y)^2 | \vec{x}[0], \vec{x}[1], ... \vec{x}[T]]
     \label{eq:vec_v}
\end{align}

In our discussion, let us assume that the expectation of $\hat{\vec{y}}$ is the
same as the true target $\vec{y}$.
\begin{align}
  E[\widehat{\vec{y}}] = \vec{y}  \label{eq:expectation_of_y}
\end{align}

Now, let us obtain the expected value of $\vec{z}$ from
\eqref{eq:vec_z} and \eqref{eq:expectation_of_y}
\begin{align}
  E[\vec{z}] = (1 - \vec{w}) \odot (\vec{x} - \vec{y})
\end{align}
Thus, the bias of $\vec{z}$ is given by:
\begin{align}
  Bias_{\vec{z}} & = E[\vec{z}] - t \nonumber \\
                 & = \vec{w} \odot (\widehat{\vec{y}} - \vec{y})
\end{align}

From \eqref{eq:vec_z} and \eqref{eq:vec_v}, the variance of $\vec{z}$ is given by:
\begin{align}
  Var_{\vec{z}} = \vec{v} \odot w \odot w.
\end{align}
The mean squared error of the $i$-th element of the $\vec{z}$ is given by:
\begin{align}
  MSE_{\vec{z_i}} & = Bias_{\vec{z_i}}^2 + Var_{\vec{z_i}} \nonumber \\
                 & =  \left[(\vec{x}_i - \vec{y}_i) ^ 2
                    + \vec{v}_i^2 \right] \vec{w}_i^{2}
                    - 2 (\vec{x}_i - \vec{y}_i)^{2}  \vec{w}_i
                    + (\vec{x}_i - \vec{y}_i)^{2}
 \label{eq:mse_z_i}
\end{align}
Since the above equation \eqref{eq:mse_z_i} is a quadratic equation with
respect to $\vec{w}_i$, we may find that the minimum value of $MSE_{\vec{z_i}}$
is obtained when $\vec{w}_i$ has the following value:
\begin{align}
  \vec{w}_i = \frac{(\vec{x}_i - \vec{y}_i)^2}{(\vec{x}_i - \vec{y}_i)^2 + \vec{v}_i}
\end{align}

Thus, the final form of the estimated vector $\widehat{\vec{z}}$ becomes:
\begin{align}
  \widehat{\vec{z}} = \widehat{\vec{w}} \odot \widehat{\vec{y}}
    \ + (1 - \widehat{\vec{w}}) \odot \vec{x}
\end{align}
where
\begin{align}
  \widehat{\vec{w}} = {(\vec{x} - \vec{y})^{\odot 2}}
    \left[(\vec{x} - \vec{y})^{\odot 2}+ \vec{v}\right]^{\odot -1}.
\end{align}

Sequence enhancement

The feature enhancement is to map a sequence of corrupt features $\mathbf{X}$.
\begin{subequations}
  \begin{align}
    \mathbf{X} = \left(\vec{x}[0], \vec{x}[1], \cdots, \vec{x}[M-1] \right)  \\
    \mathbf{Y} = \left(\vec{y}[0], \vec{y}[1], \cdots, \vec{y}[M-1] \right) 
  \end{align}
\end{subequations}

\subsubsection*{Acknowledgments}

\section*{References}

References follow the acknowledgments. Use unnumbered first-level
heading for the references. Any choice of citation style is acceptable
as long as you are consistent. It is permissible to reduce the font
size to \verb+small+ (9 point) when listing the references. {\bf
  Remember that you can go over 8 pages as long as the subsequent ones contain
  \emph{only} cited references.}
\medskip

\small

\clearpage
\newpage
\bibliographystyle{plain}
\bibliography{../../common_bib_file/common_bib_file}

\end{document}