### 2019-05-04 none_feedback_sequence_classifier/doc/main.tex
Updates the paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
%\usepackage[nonatbib]{nips_2017}

\usepackage{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{tikz,placeins}
\usepackage{subcaption}
\usepackage{algpseudocode}
%\usepackage[caption = false]{subfig}
\usepackage{amssymb,amsmath,bm}
\usetikzlibrary{arrows,decorations.markings}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\large]

\graphicspath{{../figures/}}

\title{A state of the art end-to-end speech recognition algorithm with
a homogenous structure.}

%\renewcommand{\vec}[1]{\mathbf{#1}}

\newcommand{\chanwcom}{Chanwoo Kim}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.85}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  David S.~Hippocampus\thanks{Use footnote for providing further
    information about author (webpage, alternative
    address)---\emph{not} for acknowledging funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
TODO(chanw.com) Revise the abstract.
This paper proposes a new end-to-end speech recognition system referred to
as Recurrent neural network Sequence Classification (RSC). Recently, end-to-end speech
recognition systems have been gaining more attention from researchers and
several different structures have been proposed. Typical examples  
include Connectionist Temporal Classification (CTC), Recurrent Neural Network
  Transducer (RNN-T), and Sequence-to-Sequence Modeling using the attention
mechanism. The CTC-based approaches are using the CTC loss assuming the
conditional independence property. Although the structure of CTC is simpler
than other end-to-end systems, the performance is worse due to this conditional
independence assumption. RNN-T and attention-based approaches have distinct
components such as the prediction network  and encoder in RNN-T, and
the encoder, the decoder, and the attention layer in the attention-based model.
Compared to these models, FSC has a homogeneous structure from the bottom to 
the top layer. LSTM and max-pooling layer are repeated followed by the top
softmax layer. The embedded softmax output is fed back as input of each RNN
layer. The training starts with flat initialization and alignment is made
after every epoch. The loss is the Cross Entropy (CE) loss using this 
alignment result. In our experimental results, FSC algorithm has shown 
better performance than more complicated attention-based end-to-end speech
recognition system. Another major advantage is training is significantly
faster.
\end{abstract}

\section{Introduction}
Recently, there has been tremendous improvements in speech recognition
systems fueled by advances in deep neural networks
\cite{
  Yu2013FeatureLearningDNN, 
  g_hinton_ieee_signal_processing_mag_2012,
  Seltzer2013DNNAurora4, 
  t_sainath_taslp_2017_00,
  t_sainath_book_chapter_2017_00,
  v_vanhoucke_nips_workshop_2011_00}.

TODO(chanw.com) Talk about the AI speakers.

TODO(chanw.com) Talk about the advancements of the end-to-end systems.

Recently, it has been frequently observed that if 
sequence-to-sequence end-to-end ASR systems are trained on sufficiently 
large amounts of acoustic training data, they can outperform
conventional HMM-DNN/RNN hybrid systems \cite{c_chiu_icassp_2018_00, 
c_kim_interspeech_2019_00}.

Recently, we observed that training with large-scale noisy data generated
by a \textit{Room Simulator} \cite{C_Kim_INTERSPEECH_2017_1}
improves speech recognition accuracy dramatically.
This system has been successfully employed
for training acoustic models for Google Home or Google voice
search \cite{C_Kim_INTERSPEECH_2017_1}.

\section{Review on sequence-to-sequence speech recognition algorithms}

In this section, we review well-known {\it sequence-to-sequence} algorithms
used in speech recognition.
\cite{
j_chorowski_nips_2015_00, 
a_graves_corr_2012_00, 
y_he_icassp_2019_00,
r_prabhavalkar_interspeech_2017_00}. 
A sequence-to-sequence
model maps a sequence of input acoustic features into a sequence 
of graphemes or words \cite{r_prabhavalkar_interspeech_2017_00}. 
We denote a sequence of input acoustic feature vectors by 
$\mathcal{X}_{0}^{M-1}$ and a sequence of target labels by 
$\mathcal{Y}_{0}^{L-1}$:
\begin{subequations}
  \begin{align}
    \mathcal{X}_{0}^{}  =  
      \left\{\vec{x}[0], \, \vec{x}[1], 
        \, \cdots, \, \vec{x}[M-1] \right\},  \\
    \mathcal{Y}_{0}^{}  =  
      \left\{y_0, \, y_1, \,  \cdots, \, y_{L-1} \right\},
  \end{align}
\end{subequations}
where $M$ is the number of frames in the input feature sequence and
$L$ is the number of labels in the output target sequence. In Fig. 
\ref{fig:ctc_diagram} $\sim$ \ref{fig:attention_diagram}, we show
block diagrams of CTC \cite{a_graves_icml_2006_00}, 
RNN-T \cite{a_graves_corr_2012_00, a_graves_icassp_2013_00}, and 
the attention-based model 
\cite{
j_chorowski_nips_2015_00,
w_chan_icassp_2016_00} respectively. CTC in Fig. \ref{fig:ctc_diagram} 
and RNN-T in Fig. \ref{fig:rnn_t_diagram} are {\it frame-synchronous}, 
which means that the ASR system generates the output target for each 
input frame.

\subsection{Connectionist Temporal Classification}

Fig. \ref{fig:ctc_diagram} shows the entire structure of the 
Connectionist Temporal Classification (CTC) 
\cite{a_graves_icml_2006_00}. In CTC, we use the following CTC loss
\cite{a_graves_icml_2006_00, y_he_icassp_2019_00}:
\begin{align}
  P\left(\mathcal{Y}_{0}^{L-1} | \mathcal{X}_{0}^{M-1}\right) = 
    \sum_{\mathcal{Z}_{0}^{M-1} \in 
      \mathcal{A}_{CTC} 
        \left(\mathcal{X}_{0}^{M-1}, \mathcal{Y}_{0}^{L-1} \right)}
          \Pi_{m=0}^{M-1}
            P\left(\vec{z}[m] | \mathcal{X}_{0}^{M-1} \right),
\end{align}
where $\mathcal{A}_{CTC} \left(\mathcal{X}_{0}^{M-1}, 
\mathcal{Y}_{0}^{L-1} \right)$ correspond to frame-level alignments
of length $M$ such that removing blanks and repeated symbols from 
$\mathcal{Z}_{0}^{M-1}$ yields $\mathcal{Y}_{0}^{L-1}$.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.35\textwidth}
    \centering
    \resizebox{23mm}{!}{
      \input{../figures/ctc_diagram.tex}
    } 
    \caption {
      Connectionist Temporal Classification (CTC) Model.
      \label{fig:ctc_diagram}
    }
  \end{subfigure}
  \begin{subfigure}[b]{0.60\textwidth}
    \centering
    \resizebox{78mm}{!}{
      \input{../figures/rnn_t_diagram.tex}
    }
    \caption {
      RNN-T Model.
      \label{fig:rnn_t_diagram}
    }
  \end{subfigure}

  \begin{subfigure}[b]{0.55\textwidth}
    \centering
    \resizebox{53mm}{!}{
      \input{../figures/attention_diagram.tex}
    }
    \caption {
      Attention-based Model.
      \label{fig:attention_diagram}
    }
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \resizebox{26mm}{!}{
      \input{../figures/fsc_diagram.tex}
    }
    \caption {
      Feedback Sequence Classifier (FSC) Model.
      \label{fig:fsc_diagram}
    }
  \end{subfigure}
  \caption{
    Comparison of block diagrams of different sequence-to-sequence speech
    recognition approaches.
    The proposed Feedback Sequence Classifier (FSC) is shown in 
    Fig. \ref{fig:fsc_diagram}.
  }

\end{figure}

\section{Feedback Sequence Classifier}
\label{sec:feedback_sequence_classifier}

\subsection{Neural Network Structure}
Fig. \ref{fig:fsc_diagram} shows the entire structure of the FSC. 
As shown in this figure, LSTM layers and max-pool layers
are interleaved from the bottom layer up to the third max-pool layer.

In Fig. \ref{fig:fsc_diagram}, speech features $\vec{x}[m]$ are sampled 
at every 10 {\it ms}, which is the usual speech feature frame rate  
in speech recognition
\cite{x_huang_prentice_hall_2001_00, c_kim_taslp_2016_00},
Using three two-to-one max-pool layers, $\vec{z}[p]$ has a frame period 
of 80 {\it ms}.  
The relationship between the temporal index $p$ in $\vec{z}[p]$ and 
$m$ in $\vec{x}[m]$ is given by: 
\begin{align}
  p = \lceil m / 8 \rceil \label{eq:def_p}
\end{align}

The reason for using the one dimensional max-pool layer
is to make the neural network output have a similar rate to the 
that of each {\it phone} in utterances. Note that a {\it phone} 
is any distinct speech sound serving as a phonetic unit.

It has been observed that the average phone duration is between 50 {\it ms}
\textasciitilde 100 {\it ms} \cite{x_wang_icslp_1996_00, z_bartosz_ltc_2009_00}.

\subsection{Training of the Feedback Sequence Classifier}
In this section, we describe how to train FSC. The first step is making an
alignment using the N-best alignment. The second step is updating the
neural network parameters using the Cross Entropy (CE) criterion.

\subsection{Training procedure}

The first step is finding the optimal frame-synchronous sequence
$\mathcal{Z}$

\begin{algorithmic}
  \State {$\mathcal{Z} = A(\mathcal{X} | \Theta)$}
  \State {Updates the model.}
\end{algorithmic}

\subsection{Viterbi alignment and N-best alignment}

In conventional frame-wise CE-training, Viterbi alignment 
(a.k.a forced alignment) has been performed to obtain the frame-level 
acoustic unit  boundaries TODO(chanwcom). Even though the Viterbi algorithm
has advantages in simplicity and efficiency, it is based on
the conditional independence assumption.

In TFSC, we propose the following N-best alignment approach rather than
the Viterbi alignment algorithm to find the alignment information.

\begin{algorithmic}

\For {$m = 0, . . . , M-1$}
	\For {$l = 0, . . . , N_b-1$}
		\State {$\pi^{(l)}[m]$}
	\EndFor
\EndFor

\If {$i\geq maxval$}
    \State $i\gets 0$
\Else
    \If {$i+k\leq maxval$}
        \State $i\gets i+k$
    \EndIf
\EndIf
\label{algorithm:n_best_alignment}
\end{algorithmic}

\section{Experimental Results}
\label{sec:experimental_results}

\subsubsection*{Acknowledgments}

\section*{References}

References follow the acknowledgments. Use unnumbered first-level
heading for the references. Any choice of citation style is acceptable
as long as you are consistent. It is permissible to reduce the font
size to \verb+small+ (9 point) when listing the references. {\bf
  Remember that you can go over 8 pages as long as the subsequent ones contain
  \emph{only} cited references.}
\medskip

\small

\clearpage
\newpage
\bibliographystyle{plain}
\bibliography{../../common_bib_file/common_bib_file}

\end{document}