### 2019-05-15 none_feedback_sequence_classifier/doc/main.tex
Updates the paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
%\usepackage[nonatbib]{nips_2017}

\usepackage{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{tikz,placeins}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage[caption = false]{subfig}
\usepackage{amssymb,amsmath,bm}
\usetikzlibrary{arrows,decorations.markings}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\large]

\graphicspath{{../figures/}}

\title{A state of the art end-to-end speech recognition algorithm with
a homogenous structure.}

%\renewcommand{\vec}[1]{\mathbf{#1}}

\newcommand{\chanwcom}{Chanwoo Kim}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.85}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  David S.~Hippocampus\thanks{Use footnote for providing further
    information about author (webpage, alternative
    address)---\emph{not} for acknowledging funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
TODO(chanw.com) Revise the abstract.
This paper proposes a new end-to-end speech recognition system referred to
as Feedback Sequence Classifier (FSC). Recently, end-to-end speech
recognition systems have been gaining more attention from researchers and
several different structures have been proposed. Typical examples  
include Connectionist Temporal Classification (CTC), Recurrent Neural Network
  Transducer (RNN-T), and Sequence-to-Sequence Modeling using the attention
mechanism. The CTC-based approaches are using the CTC loss assuming the
conditional independence property. Although the structure of CTC is simpler
than other end-to-end systems, the performance is worse due to this conditional
independence assumption. RNN-T and attention-based approaches have distinct
components such as the prediction network  and encoder in RNN-T, and
the encoder, the decoder, and the attention layer in the attention-based model.
Compared to these models, FSC has a homogeneous structure from the bottom to 
the top layer. LSTM and max-pooling layer are repeated followed by the top
softmax layer. The embedded softmax output is fed back as input of each RNN
layer. The training starts with flat initialization and alignment is made
after every epoch. The loss is the Cross Entropy (CE) loss using this 
alignment result. In our experimental results, FSC algorithm has shown 
better performance than more complicated attention-based end-to-end speech
recognition system. Another major advantage is training is significantly
faster.
\end{abstract}

\section{Introduction}
Recently, there has been tremendous improvements in speech recognition
systems fueled by advances in deep neural networks
\cite{
  Yu2013FeatureLearningDNN, 
  g_hinton_ieee_signal_processing_mag_2012,
  Seltzer2013DNNAurora4, 
  t_sainath_taslp_2017_00,
  t_sainath_book_chapter_2017_00,
  v_vanhoucke_nips_workshop_2011_00}.
TODO(chanw.com) Describes history about starting using DNN, LSTM.

With widespread adoption of voice assistant speakers
such as Google Home, Amazon Alexa
\cite{B_Li_INTERSPEECH_2017_1, C_Kim_INTERSPEECH_2017_1},
far-field speech recognition has become very important.
Recently, approaches using data augmentation
has been gaining popularity for far-field speech recognition
\cite{R_Lippmann_icassp_1987_1,
c_kim_interspeech_2018_00, w_hartmann_interspeech_2016_00}.

TODO(chanw.com) Briefly describes CTC, RNN-T, and attention-based model.

Recently, it has been observed that if 
{\it sequence-to-sequence end-to-end} ASR systems are trained on sufficiently 
large amounts of acoustic training data, they can outperform
conventional HMM-DNN/RNN hybrid systems \cite{c_chiu_icassp_2018_00, 
c_kim_interspeech_2019_00}.
These complete end-to-end systems have started surpassing the performance of
the conventional WFST-based decoders with a improved training strategy
such as label smoothing, Minimum Word Error Rate(MWER) training, 
learning rate warming-up,
and a better choice of target unit such as Byte Pair Encoded (BPE) 
\cite {r_sennrich_acl_2016_00} subword units.

TODO(chanw.com) Briefly describes the motivation of FSC.

\section{Review on sequence-to-sequence speech recognition algorithms}

In this section, we review well-known {\it sequence-to-sequence} algorithms
used in speech recognition.
\cite{
j_chorowski_nips_2015_00, 
a_graves_corr_2012_00, 
y_he_icassp_2019_00,
r_prabhavalkar_interspeech_2017_00}. 
A {\it sequence-to-sequence}
speech recognizers maps a sequence of input acoustic features into a sequence 
of text labels  \cite{r_prabhavalkar_interspeech_2017_00}. 
Mathematically, this model is 
represented by the following transformation $\mathcal{T}$ 
\cite{a_graves_icml_2006_00}:
\begin{align}
  \mathcal{T} : \left(\mathbb{R}^d\right)^{*} \rightarrow \mathbb{V}^{*}
	\label{eq:transformation}
\end{align}
where the symbol ${*}$ denotes a sequence, $d$ is the dimension of the input 
feature. $\mathbb{V}$ in \eqref{eq:target_label} is the set of output
labels, which may be graphemes 
\cite{j_chorowski_nips_2015_00,
w_chan_icassp_2016_00},
subword units \cite{
a_zeyer_interspeech_2018_00,
c_chiu_icassp_2018_00},
 and words \cite{h_soltau_interspeech_2017_00}.
Let us denote a sequence of input acoustic feature vectors by 
$\mathcal{X}_{0}^{M-1}$ and a sequence of target labels in 
{\it one-hot vector} representation or its label-smoothed 
representation \cite{} by 
$\mathcal{Y}_{0}^{L-1}$ as shown below:
\begin{subequations}
  \begin{align}
    \mathcal{X}_{0}^{M-1}  =  
      \left\{\vec{x}[m]
          \Big| 0 \le m \le M - 1, \; \vec{x}[m] \in \mathbb{R}^d \right\},  
    \label{eq:input_seq} \\
    \mathcal{Y}_{0}^{L-1}  =  
      \left\{\vec{y}_l  \Big| 0 \le l \le L-1, \; \vec{y}_l \in \mathbb{V} \right\}, 
    \label{eq:target_label}
  \end{align}
\end{subequations}
where $M$ is the number of frames in the input feature sequence,
$L$ is the number of labels in the output target sequence,
 and $\mathbb{V}$ in \eqref{eq:target_label} is the set of output
Note that the sequence index in \eqref{eq:input_seq} is a frame
index, whereas the sequence index in \eqref{eq:target_label} is 
label index. Depending on whether the neural network generates
the inference output for every frame or not, there are following
categories:

\subsection{Label-synchronous inference}
In this class of models, the neural network generates the inference output 
$\widehat{\vec{y_l}}$ 
only when a new label is expected. The loss function 
$\mathbb{L}_{\text{ls}}\left(\mathcal{X}_{0}^{M-1}, \mathcal{Y}_{0}^{L-1}\right)$ is 
defined as follows:
   \begin{align}
     \mathbb{L}_{\text{ls}}\left(\mathcal{X}_{0}^{M-1}, \mathcal{Y}_{0}^{L-1}\right) 
        = \sum_{l=0}^{L-1}  {\vec{y}_l}^{\;t}  \log\left( 
            \circ \, \widehat{\vec{y_l}} \right)  
          \label{eq:label_synchronous_ce_loss} 
   \end{align}
where $\log \left(\circ \right)$ and $^{t}$ denote element-wise log and 
transpose operation, respectively.

\subsection{Frame-synchronous inference}
\label{sec:frame_synchronous_inference}
In this class of inference,  we obtain
the temporal boundaries of each label in \eqref{eq:target_label}.
In the frame-wise Cross Entropy (CE) training 
\cite{g_hinton_ieee_signal_processing_mag_2012_00, 
C_Kim_INTERSPEECH_2017_1, 
B_Li_INTERSPEECH_2017_1}, {\it Viterbi alignment}  
\cite{
x_huang_prentice_hall_2001_00,
l_r_rabiner_proceedings_of_ieee_1989_00} is usually employed. 
The FSC algorithm presented in this paper also obtains
label boundaries using the {\it N-best alignment} which will
be explained in Sec. \ref{sec:n_best_alignment}.

Using these alignment algorithms, 
we obtain corresponding {\it frame-synchronous} label sequences 
$\vec{z}[m] \in \mathbb{V}$ for each frame index $m$. This
latent label sequences are often called a {\it path}.
  \begin{align}
    \mathcal{Z}_{0}^{M-1}  =  
      \left\{\vec{z}[m]  \Big| 0 \le m \le M-1, \; \vec{z}[m] \in \mathbb{V} \right\}, 
    \label{eq:latent_label_seq}
  \end{align}
Fig. \ref{fig:path_alignment} shows one such example. 
Such an alignment is usually called a {\it path}.

\begin{align}
   \mathbb{L}_{\text{fs}}\left(\mathcal{X}_{0}^{M-1}, \mathcal{Z}_{0}^{M-1}\right) 
      = \sum_{l=0}^{M-1}  {\vec{z}[m]}^{\,t}  \log\left( 
          \circ \, \widehat{\vec{z}}[m] \right)  
        \label{eq:frame_synchronous_ce_loss} 
\end{align}

 In Fig. 
\ref{fig:ctc_diagram} $\sim$ \ref{fig:attention_diagram}, we show
block diagrams of CTC \cite{a_graves_icml_2006_00}, 
RNN-T \cite{a_graves_corr_2012_00, a_graves_icassp_2013_00}, and 
the attention-based model 
\cite{
j_chorowski_nips_2015_00,
w_chan_icassp_2016_00} respectively. CTC in Fig. \ref{fig:ctc_diagram} 
and RNN-T in Fig. \ref{fig:rnn_t_diagram} are {\it frame-synchronous}, 
which means that the ASR system generates the output target for each 
input frame.

\subsubsection{Connectionist Temporal Classification}
\label{sec:ctc}

Fig. \ref{fig:ctc_diagram} shows the entire structure of the 
Connectionist Temporal Classification (CTC) 
\cite{a_graves_icml_2006_00}. In CTC, we use the following CTC loss
\cite{a_graves_icml_2006_00, y_he_icassp_2019_00}:
\begin{align}
  P\left(\mathcal{Y}_{0}^{L-1} | \mathcal{X}_{0}^{M-1}\right) = 
    \sum_{\mathcal{Z}_{0}^{M-1} \in 
      \mathcal{A}_{CTC} 
        \left(\mathcal{X}_{0}^{M-1}, \mathcal{Y}_{0}^{L-1} \right)}
          \Pi_{m=0}^{M-1}
            P\left(\vec{z}[m] | \mathcal{X}_{0}^{M-1} \right),
\end{align}
where $\mathcal{A}_{CTC} \left(\mathcal{X}_{0}^{M-1}, 
\mathcal{Y}_{0}^{L-1} \right)$ correspond to frame-level alignments
of length $M$ such that removing blanks and repeated symbols from 
$\mathcal{Z}_{0}^{M-1}$ yields $\mathcal{Y}_{0}^{L-1}$.

To obtain derivatives required for back-propagation, CTC employs 
the forward-backward algorithm.

\subsubsection{RNN-Transducer}
\label{sec:rnn_t}

Fig. \ref{fig:rnn_t_diagram} shows the entire structure of the  
RNN-T structure. The loss function used in RNN-T is given as follows:
\begin{align}
  P\left(\mathcal{Y}_{0}^{L-1} | \mathcal{X}_{0}^{M-1}\right) = 
    \sum_{\mathcal{Z}_{0}^{M-1} \in 
      \mathcal{A}_{CTC} 
        \left(\mathcal{X}_{0}^{M-1}, \mathcal{Y}_{0}^{L-1} \right)}
          \Pi_{m=0}^{M-1}
            P\left(\vec{z}[m] | \mathcal{X}_{0}^{M-1} \right),
\end{align}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.35\textwidth}
    \centering
    \resizebox{23mm}{!}{
      \input{../figures/ctc_diagram.tex}
    } 
    \caption {
      Connectionist Temporal Classification (CTC) Model.
      \label{fig:ctc_diagram}
    }
  \end{subfigure}
  \begin{subfigure}[b]{0.60\textwidth}
    \centering
    \resizebox{78mm}{!}{
      \input{../figures/rnn_t_diagram.tex}
    }
    \caption {
      RNN-T Model.
      \label{fig:rnn_t_diagram}
    }
  \end{subfigure}

  \begin{subfigure}[b]{0.55\textwidth}
    \centering
    \resizebox{53mm}{!}{
      \input{../figures/attention_diagram.tex}
    }
    \caption {
      Attention-based Model.
      \label{fig:attention_diagram}
    }
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \resizebox{26mm}{!}{
      \input{../figures/fsc_diagram.tex}
    }
    \caption {
      Feedback Sequence Classifier (FSC) Model.
      \label{fig:fsc_diagram}
    }
  \end{subfigure}
  \caption{
    Comparison of block diagrams of different sequence-to-sequence speech
    recognition approaches.
    The proposed Feedback Sequence Classifier (FSC) is shown in 
    Fig. \ref{fig:fsc_diagram}.
  }

\end{figure}

\section{Feedback Sequence Classifier}
\label{sec:feedback_sequence_classifier}

\subsection{Neural Network Structure}
\label{sec:neural_network_structure}
Fig. \ref{fig:fsc_diagram} shows the entire structure of the FSC. 
Speech features $\vec{x}[m]$ are sampled 
at every 10 {\it ms}, which is the usual frame period in speech 
\cite{x_huang_prentice_hall_2001_00, c_kim_taslp_2016_00}.
As shown in this figure, LSTM layers and max-pool layers
are interleaved from the bottom layer up to the third max-pool layer.
Using three 2:1 max-pool layers, the softmax output $\vec{z}[u]$ has 
a frame period of 80 {\it ms}.  Thus, the relationship between the original frame 
index $m$ and the temporal sub-sampled index $u$ is given by \eqref{eq:def_u_U}.
The relationship between the length of the softmax output sequence $U$
and the length of the feature sequence $M$ is also given in the same way:
	\begin{align}
		u = \lceil m / 8 \rceil,  \; U = \lceil M / 8 \rceil \label{eq:def_u_U}.
	\end{align}

FSC performs inference in a frame-synchronous way as described in 
Sec. \ref{sec:frame_synchronous_inference}. 
However, due to this temporal sub-sampling, the latent label sequence 
in \eqref{eq:latent_label_seq} is slightly modified as shown below:
\begin{align}
  \mathcal{Z}_{0}^{U-1}  =  
    \left\{\vec{z}[u]  \Big| 0 \le m \le U-1, \; \vec{z}[u] \in \mathbb{V} \right\}.
  \label{eq:subsampled_latent_label_seq}
\end{align}

The loss function 
$\mathbb{L}_{\text{fsc}}\left(\mathcal{X}_{0}^{M-1}, \mathcal{Z}_{0}^{M-1}\right)$ 
is also modified from the original one in 
\eqref{eq:frame_synchronous_ce_loss}:
\begin{align}
	\mathbb{L}_{\text{fsc}}
		\left(\mathcal{X}_{0}^{M-1}, \mathcal{Z}_{0}^{U-1}\right)
      = \sum_{u=0}^{U-1}  {\vec{z}[u]}^{\,t}  \log\left( 
          \circ \, \widehat{\vec{z}}[u] \right)  
        \label{eq:fsc_ce_loss} 
\end{align}
The reason for performing the 8:1 temporal sub-sampling in \eqref{eq:def_u_U} is to 
to make the period of the softmax output $\widehat{\vec{z}}[u]$ in \label{eq:fsc_ce_loss} 
be similar to the shortest BPE duration. We assume that the {\it sequence-to-sequence}
neural network models will be trained most efficiently when the period of neural 
network output roughly matches the period of the target. In addition, this
output periods must be equal to larger than the shortest target period to prevent
missing some targets.
The shortest duration of each BPE is similar to the average length of {\it phones}, 
which is any distinct speech sound serving as a phonetic unit.
It has been observed that the average phone duration is between 50 {\it ms}
\textasciitilde 100 {\it ms} \cite{x_wang_icslp_1996_00, z_bartosz_ltc_2009_00}.
Under this rationale, we chose the temporal sub-sampling factor of 8:1. 
We also observe this sub-sampling factor is optimal in our experiments as well.

\subsection{Target units of Feedback Sequence Classifier}
\label{sec:target_units}

In FSC, we use subword units created by BPE \cite{r_sennrich_acl_2016_00} for
the output vocabulary $\mathbb{V}$ in 
\eqref{eq:transformation} and \eqref{eq:subsampled_latent_label_seq}.

When creating the latent frame-wise label sequence $\mathcal{Z}_{0}^{U-1}$ 
in \eqref{eq:subsampled_latent_label_seq}, we distinguish the onset frame
from other frames.  For example, if we have a BPE unit
("{\it an}") in the vocabulary $\mathbb{V}$, then we define another
target ("{\it an$_o$}") which marks the on-set frame of this BPE
interval. Thus, when the number of BPE units is $B$, the total
number of vocabularies in $\mathbb{V}$ becomes $2B$.
We refer this approach to as {\it BPE with onset marking}.

The reason for using this approach is two-folds. First, since FSC
generates an inference output for each frame, if the same
BPE symbol repeatedly appears, it is difficult to tell whether 
this BPE is really expected to appear multiple times or whether 
the a single BPE is continuing. For example, when we have 
a sequence \{"{\it an}", "{\it an} "\}
as the inference output, it is difficult to tell whether it should be a single
"{\it an}" or two "{\it an}"s. With the onset marking,
a decoder needs to print out the inference outputs with onset marking. 
The blank symbol in CTC also plays a similar role. However this blank symbol
is a common symbol following any labels. In the FSC algorithm, the information
given by the feedback paths is crucial in performance. In this aspect,
blank symbol is not very useful for this feedback structure. 
Second, we have frequently observed that the on-set portion of sound
has a very important role in human auditory perception 
\cite{C_Kim_INTERSPEECH_2014_2}.

\subsection{Training of the Feedback Sequence Classifier}
In this section, we describe how to train FSC.  Algorithm 
\ref{algo:fsc_training} shows how FSC is trained using the 
N-best alignment algorithm. N-best algorithm will be described 
in Sec. \ref{sec:n_best_alignment} in detail.
In Algorithm \ref{algo:fsc_training}  
As mentioned in Sec. 
\ref{sec:frame_synchronous_inference}, FSC performs inference
in a frame-synchronous way.
The first step is making an 
In this section, we describe how to train FSC. The first step is 
is to make an alignment using the N-best alignment. The second step is updating the
neural network parameters using the Cross Entropy (CE) criterion.

The first step is finding the optimal frame-synchronous sequence
$\mathcal{Z}_0^{M-1}$.   
$\mathcal{Z}$

\begin{algorithm}
  \caption{Training procedure for the FSC algorithm}
  \label{algo:fsc_training}
  \begin{algorithmic}
    \State $\Theta^{(0)} \leftarrow$ initialize\_parameters()
    \For {epoch index $i = 0, . . . , I-1$}
      \For {each batch $\left\{\mathcal{X}_{0}^{M-1}  \right\}, 
                        \left\{\mathcal{Y}_{0}^{L-1}  \right\}$}, 
        \State $\left\{ {{\mathcal{Z}^{(i)}}_0^{U-1}} \right\}
          \leftarrow
          \text{n\_best\_align}
          \left(
            \left\{X_{0}^{M-1} \right\} \Big| \Theta^{(i)}
          \right)$

        \State {$\Theta^{(i)} \leftarrow$ update\_parameter 
          $\left(
            \left\{\mathcal{X}_{0}^{M-1} \right\},			
            \left\{{\mathcal{Z}^{(i)}}_{0}^{U-1}\right\}
          \right)$}

    \EndFor
    \State $\Theta^{(i+1)} := \Theta^{(i)}$
  \EndFor
  \label{algorithm:fsc_training}
  \end{algorithmic}
\end{algorithm}

\subsection{N-best alignment}
\label{sec:n_best_alignment}

\begin{figure}
  \centering
    \resizebox{100mm}{!}{
      \input{../figures/path_alignment.tex}
    } 
    \caption {
      An example of a path $\mathcal{Z}_{0}^{M-1}$ and the path movement 
      constraint shown by dashed arrows $P \rightarrow Q$ and $P \rightarrow R$. 
      \label{fig:path_alignment}
    }
\end{figure}

In CTC in Sec. \ref{sec:ctc} and in RNN-T in Sec. \ref{sec:rnn_t}, the 
forward-backward algorithm is employed to update parameters. 
For example in CTC, the forward-backward algorithm is based on the
following assumption:
\begin{align}
  P \left(  \Big| \right) = \Pi
\end{align}

In conventional frame-wise CE-training, Viterbi alignment 
(a.k.a forced alignment) has been performed to obtain the frame-level 
acoustic unit  boundaries TODO(chanwcom). Even though the Viterbi algorithm
has advantages in simplicity and efficiency, it is based on
the following conditional independence assumption.

In TFSC, we propose the following N-best alignment approach rather than
the Viterbi alignment algorithm to find the alignment information.

\begin{algorithm}
  \caption{N-best alignment algorithm}
  \begin{algorithmic}

  \Require {$\mathcal{X}_0^{M}$}
  \Require {$\mathcal{Y}_0^{L}$}

  \For {$m = 0, . . . , M-1$}
    \For {$l = 0, . . . , N_b-1$}
      \State {$\pi^{(l)}[m]$}
    \EndFor
  \EndFor

  \If {$i\geq maxval$}
      \State $i\gets 0$
  \Else
      \If {$i+k\leq maxval$}
          \State $i\gets i+k$
      \EndIf
  \EndIf
  \label{algorithm:n_best_alignment}
  \end{algorithmic}
\end{algorithm}

\section{Experimental Results}
\label{sec:experimental_results}

\subsubsection*{Acknowledgments}

\small

\clearpage
\newpage
\bibliographystyle{plain}
\bibliography{../../common_bib_file/common_bib_file}

\end{document}