### 2019-07-02 interspeech_2019_training_infra/doc/main.tex
Updates papers

% Template for ICASSP-2019 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass[a4paper]{article}

\usepackage{INTERSPEECH2019}
\usepackage{amsmath,graphicx}
\usepackage{amssymb,amsmath,bm}
\usepackage{textcomp}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{multirow}
%\usepackage{subcaption}
\usepackage[justification=centering]{caption}
\usepackage[caption = false]{subfig}

% Custom settings
\usepackage{tikz,placeins}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\large]

\graphicspath{{../figures/}}

\newcommand{\specialcell}[2][c]{%
     \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newcommand{\chanwcom}{{C. Kim}}
\renewcommand{\topfraction}{1.0}
\renewcommand{\bottomfraction}{1.0}
\renewcommand{\textfraction}{0.01}
\renewcommand{\floatpagefraction}{1.0}
\renewcommand{\dbltopfraction}{1.0}
\renewcommand{\floatpagefraction}{1.0}  % require fuller float pages
\renewcommand{\dblfloatpagefraction}{1.0} % require fuller float pages

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

% Title.
% ------
\title{End-to-end Training of a Large Vocabulary \\ End-to-end Speech Recognition System}
%\title{End-to-end training framework with on-the-fly data augmentation for building a large vocabulary end-to-end speech secognition system}
%
% Single address.
% ---------------
\name{Chanwoo Kim, Sungsoo Kim, Kwangyoun Kim, Mehul Kumar,
Jiyeon Kim, Kyungmin Lee, Changwoo Han, Abhinav Garg, Eunhyang Kim, Minkyoo
Shin, Shatrughan Singh, Larry Heck, Dhananjaya Gowda
\thanks{Thanks to Samsung
Electronics for funding this research. The authors are thankful to  
Executive Vice President Seunghwan Cho and members of Speech Processing Lab
 at Samsung Research.}}

\address{Samsung Research }
\email{\{chanw.com, ss216.kim, ky85.kim, mehul3.kumar, jstacey7.kim,
k.m.lee, cw1105.han, abhinav.garg, sc.ehkim.jin, mk0211.shin, shatrughan.s, 
larry.h, d.gowda\}@samsung.com }
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
\ninept
%
\maketitle
%
\begin{abstract}
In this paper, we present an end-to-end training framework and strategies for 
building state-of-the-art end-to-end (E2E) speech recognition systems.
Our training system utilizes a cluster of Central Processing Units 
  (CPUs) and Graphics Processing Units (GPUs).
The entire data reading, large scale data augmentation,
neural network parameter updates are all performed on-the-fly. 
  We use vocal tract length perturbation and an acoustic simulator 
for data augmentation. The processed features and labels are sent 
to the GPU cluster. Horovod {\tt allreduce} approach is employed to train
neural network parameters.
We evaluated the effectiveness of our system on the standard 
Librispeech corpus \cite{v_panayotov_icassp_2015_00} and 
  10,000-hr anonymized Bixby English dataset. 
Our end-to-end speech
recognition system built using this training infrastructure showed 
  2.85 \% WER on {\tt test-clean} for the LibriSpeech corpus 
 after shallow fusion with a recurrent neural network (RNN) language model (LM).
 For English Bixby open domain test set,
we obtained a WER of 7.92 \% using 
a bidirectional full attention (BFA) E2E model after shallow fusion with an RNN-LM.
When the monotonic chunckwise attention (MoCha) based approach is employed for 
streaming speech recognition, we obtained a WER of 9.95 \% 
  on the same Bixby open domain test set.
\end{abstract}
%
%\begin{keywords}
  \noindent{\bf Index Terms}: end-to-end speech recognition,
distributed training, example server, data augmentation, acoustic simulation
%\end{keywords}
%
%
\section{Introduction}
In recent years, deep learning techniques have significantly 
improved speech recognition accuracy \cite{Seltzer2013DNNAurora4, 
Yu2013FeatureLearningDNN, V_Vanhoucke_Deep_Learning_NIPS_Workshop_2011,
G_Hinton_IEEE_Signal_Process_Mag_2012,
T_Sainath_IEEETran_2017_1, T_Sainath_Book_Chapter_2017_1}.
This improvement has come about from the shift from Gaussian Mixture Model
(GMM) to the Feed-Forward Deep Neural Networks (FF-DNNs), FF-DNNs
to Recurrent Neural Network (RNN) and in particular the Long Short-Term Memory
(LSTM) networks. 
Thanks to these advances, voice assistant devices such as Google Home
\cite{C_Kim_INTERSPEECH_2017_1, B_Li_INTERSPEECH_2017_1}
or Amazon Alexa or Samsung Bixby are being used at many homes and on personal devices.

Recently there has been increasing interest in switching
from the conventional Weighted Finite State Transducer (WFST)
based decoder using an Acoustic Model (AM) and a Language Model (LM)
to a complete end-to-end all-neural speech recognition systems 
\cite{w_chan_icassp_2016_00, r_prabhavalkar_interspeech_2017_00, 
j_chorowski_nips_2015_00}. 
These complete end-to-end systems have started surpassing the performance of
the conventional WFST-based decoders with a very large training 
database, a better choice of target unit such as Byte Pair Encoded (BPE) 
subword units, and an improved training methodology such as Minimum
Word Error Rate (MWER) training.

Another important aspect of these recent improvements in speech
recognition performance is data. 
To build high performance speech recognition systems for
conversational speech, we need to use
a large amount of speech data covering various domains. In
\cite{h_soltau_interspeech_2017_00}, it has been shown
that we need a very large training set ($\sim$125,000 hours of 
semi-supervised speech data) to achieve
high speech recognition accuracy for difficult tasks like 
video captioning. To train neural networks using such large 
amounts of speech data, we usually need 
multiple Central Processing Units (CPUs) or Graphics
Processing Units (GPUs) or GPU clusters 
\cite{E_Variani_INTERSPEECH_2017_01, p_goyal_arxiv_2018_00}.

With widespread adoption of voice assistant speakers, far-field
speech recognition has become very important.
In far-field speech recognition, the impacts of reverberation and noise 
are much larger than those in near-field cases. 
Traditional approaches to far-field
speech recognition include noise robust feature extraction algorithms
\cite{C_Kim_IEEETran_2016_1, U_H_Yapanel_SpeechComm_2008},
or multi-microphone approaches
\cite{T_Nekatani_ICASSP_2017_1, T_Higuchi_ICASSP_2016_1,
H_Erdogan_INTERSPEECH_2016_1,
C_Kim_INTERSPEECH_2010_1, C_Kim_ICASSP_2012_2}.
More recently, approaches using data augmentation
has been gaining popularity for far-field speech recognition
\cite{R_Lippmann_icassp_1987_1,
c_kim_interspeech_2018_00, w_hartmann_interspeech_2016_00}. 
An ``acoustic simulator"  
\cite{C_Kim_INTERSPEECH_2017_1, c_kim_interspeech_2018_00}
is used to generate simulated speech utterances for
millions of different room dimensions, a wide distribution
of reverberation time and signal-to-noise ratio. In a similar 
spirit, Vocal Tract Length Perturbation (VTLP) has been proposed 
\cite{n_jaitly_icml_workshop_2013_00} to tackle the speaker variability issue.
As shown in our recent paper \cite{c_kim_interspeech_2019_00}, 
VTLP is especially useful when
the speaker variability in the training database is not sufficient. 
For these kinds of data augmentation, processing on CPUs is more
desirable than processing on GPUs. Due to this, we have proposed
an end-to-end training approach using Example Servers (ES)
 and workers. Example servers are typically run on the CPU
cluster performing data reading, data augmentation, and feature extraction
\cite{c_kim_interspeech_2018_00}.

%
%
\begin{figure*}[!tbp]
  \begin{center}
    \resizebox{\textwidth}{!}{\input{../figures/end_to_end_training_system.tex}}
      \caption { The Samsung Research end-to-end training framework for building an end-to-end
	speech recognition system with multi CPU-GPU clusters and on-the-fly data
	processing and augmentation pipeline.
     }   
     \label{fig:end_to_end_training_system}
  \end{center}
\vspace{-5mm}
\end{figure*}
%
%

In this paper, we describe the structure of our end-to-end training system
to train an end-to-end speech recognition system. This training system
has several advantages over previous systems described in
\cite{c_kim_interspeech_2018_00}. First, instead of using the {\tt QueueRunner},
we use a more efficient data queue using {\tt tf.data} in Tensorflow 
\cite{m_abadi_usenix_2016}.  Second, instead
of pre-calculating information about room configurations and 
room impulse responses in the acoustic simulator, these are
calculated on-the-fly. Thus, the entire training system runs on
-the-fly. Additionally, instead of using the parameter server-worker
structure, we use an {\tt allreduce} approach implemented in the Horovod
\cite{a_sergeev_arxiv_2018_00} distributed training framework, which
has been shown to be more efficient. The system described in 
\cite{E_Variani_INTERSPEECH_2017_01}, is designed
to train the acoustic model part of the speech recognition system where
as our training system trains the complete end-to-end speech recognition system.

The rest of the paper is organized as follows: We describe the entire
training system structure in detail in Sec. \ref{sec:training_cpu_gpu}.
The structure of the end-to-end speech recognition system 
is described in Sec. \ref{sec:end_to_end_speech_recognition}. 
Experimental results that demonstrates the effectiveness of our speech
recognition system is presented in Sec. \ref{sec:experimental_results}. 
We conclude in Sec.  \ref{sec:conclusions}.
%
\section{Overall structure of the end-to-end speech recognition}
\label{sec:training_cpu_gpu}
\vspace{-1mm}

In this section, we describe the overall structure of our 
end-to-end training system.  Fig. \ref{fig:end_to_end_training_system}
shows how the entire system is structured. Our system consists
of a cluster of CPUs and a cluster of GPUs. Each GPU node of 
the GPU cluster has eight nVidia\texttrademark P-40 GPUs and two 
Intel E5-2690 v4 CPUs. Each of these CPUs has 14 cores.
The large box on the left hand side of Fig. \ref{fig:end_to_end_training_system}
denoted ``GPU cluster'' shows a typical GPU node with $N$ GPUs.
The large box on the right shows a ``CPU cluster'' of $M$ CPUs,
each running an independent data pipeline.

\vspace{-1mm}
\subsection{Training job launch}
The main process of the training system runs on one of CPU cores 
of the GPU cluster. This CPU portion of the GPU node is represented
as a box in the right hand side of the GPU node box.
When the training job starts, this main training process launches
multiple example server jobs on the CPU cluster using the 
{\tt IBM Platform LSF} \cite{ibm_spectrum_lsf_2010}. In Fig. 
\ref{fig:end_to_end_training_system}, this launching process is represented
by a dashed arrow from the CPU portion of the GPU node to the 
CPU cluster.

\subsection{Data reading using an example queue}
In the CPU cluster, each CPU runs one example server
which reads speech utterance and transcript data from sharded 
{\tt TFRecords} defined in Tensorflow \cite{m_abadi_usenix_2016}.
The TFRecord format is a simple format in Tensorflow for storing 
a sequence of binary records. To support efficient reading using 
multiple CPUs, we use sharded {\tt TFRecords}.

To read large-scale data efficiently in parallel, we use an example queue
shown in the left side of Fig. \ref{fig:end_to_end_training_system}.  
The original speech waveform data, transcripts, and meta data are stored
in sharded {\tt TFRecords}. The data pipeline is implemented using
{\tt tf.data} in Tensorflow \cite{m_abadi_usenix_2016}, 
and contains the data augmentation and feature extraction blocks.
These {\tt tf.data} APIs are efficient in building complex pipelines
by applying a series of elementary operations. We perform data interleaving
and parallel reading using {\tt tf.contrib.data.parallel\_interleave}, 
shuffling using {\tt tf.data.Datatset.shuffle}, and padding using 
{\tt tf.data.Dataset.padded\_batch}.

\subsection{Data augmentation and feature extraction}
\label{sec:feature_extraction}
To improve robustness against speaker variability, we apply an
on-the-fly VTLP algorithm on the input waveform. The warping
factor is generated randomly for each input utterance. Unlike
conventional VTLP approaches in \cite{x_cui_taslp_2015_00,
n_jaitly_icml_workshop_2013_00}, we resynthesize the processed
speech. The purpose of doing this is to apply VTLP before applying
the acoustic simulator. One more advantage is that this
resynthesis approach enables us to use a window length 
optimal for VTLP different from that used in feature processing.
In the bilinear transformation, the relation between the input
and output discrete-time frequencies is given by: 
\begin{align}
  \omega_k'  = \omega_k + 2 \tan^{-1} \left(
                    \frac{ \left(1 - \alpha \right) \sin(\omega_k)} 
              {1 - (1 - \alpha) \cos(\omega_k) } \right)
                            \label{eq:bilinear_transformation}.
\end{align}
where $w_k= {2 \pi k}/{K}$ is the discrete-time frequency
 and $K$ is the DFT size. 
More details about our VTLP algorithm can be found
in our another paper \cite{c_kim_interspeech_2019_00}.
The acoustic simulator in Fig.~\ref{fig:end_to_end_training_system}
is similar to what we described in 
\cite{C_Kim_INTERSPEECH_2017_1, c_kim_interspeech_2018_00}.
One difference compared to our previous one in 
\cite{C_Kim_INTERSPEECH_2017_1} is that we do not pre-calculate
room impulse responses, but instead they are calculated on-the-fly.
For feature processing we use {\tt tf.data.Dataset.map} API. 
Instead of using the more conventional log-mel or 
MFCC features, we use the power mel filterbank energies,
since it shows slightly better performance \cite{c_kim_interspeech_2019_01}.

\subsection{Parameter calculation and update}
The features and the target label are sent to the GPU cluster
using the ZeroMQ \cite{zero_mq} asynchronous messaging queue.
Each example server sends these data asynchronously to the CPU portion
of the GPU node as shown in Fig \ref{fig:end_to_end_training_system}.
Using these data, neural network parameters are calculated and updated 
using an Adam optimizer and the
Horovod \cite{a_sergeev_arxiv_2018_00} {\tt allreduce} approach.

\section{Structure of the end-to-end speech recognition system}
\label{sec:end_to_end_speech_recognition}

\begin{figure}[tbp]
  \begin{center}
    \resizebox{65mm}{!}{\input{../figures/e2e_asr_block_diagram}}
      \caption {  The structure of the entire end-to-end
      speech recognition system. 
      %The LSTMs
      %in the encoder layers may be either bidirectional-LSTMs or
      %unidirectional-LSTMs. The attention may be either the full
      %attention or the MOnotonic CHunkwise Attention (MoCha)
      %\cite{c_chiu_iclr_2018_00}.
     }
     \label{fig:entire_diagram}
  \end{center}
\vspace{-5mm}
\end{figure}

We have adopted the RETURNN speech recognition system \cite{p_doetsch_icassp_2017_00,
a_zeyer_interspeech_2018_00} for training our end-to-end system with various modifications.
Some of the important modifications are: replacing the input data pipeline with our 
proposed on-the-fly example server based pipeline with support for VTLP
and acoustic simulation,
implementing the Monotonic Chunkwise Attention (MoChA) \cite{c_chiu_iclr_2018_00}
for online streaming E2E speech recognition, minimum Word Error Rate (mWER) training,
support for handling Korean language or script, our own scoring and Inverse Text 
Normalization modules, support for power mel filterbank features, Maximally Uniform Distribution (MUD) preprocessing of data, etc.

The structure of our entire end-to-end speech recognition system
with MUD processing \cite{c_kim_interspeech_2019_01} is shown 
in Fig.~\ref{fig:entire_diagram}.
 $\vec{x}[m]$ and $\vec{y}[n]$ are the input power mel 
filterbank energy vector and the output label,
respectively. $m$ is the input frame index and $n$ is the decoder output
step index. $\vec{c}[n]$ is the attention context vector calculated 
as a weighted sum of the encoder hidden state vectors denoted as $\vec{h}_{enc}[m]$.
The attention weights are computed as a softmax of energies computed as 
a function of the encoder hidden state $\vec{h}_{enc}[m]$, the decoder hidden state
$\vec{h}_{dec}[n]$, and the attention weight
feedback $\vec{\beta}[m, n]$  \cite{a_zeyer_interspeech_2018_00}.

In \cite{a_zeyer_interspeech_2018_00}, the peak value of the speech 
waveform is normalized to be one. However, since finding the peak sample value 
is not possible for online feature extraction, we do not perform this
normalization. We modified the input pipeline
so that the online feature generation can be performed. We disabled the 
clipping of feature range between -3 and 3, which is the default 
setting for the Librispeech experiment using MFCC 
features in \cite{a_zeyer_interspeech_2018_00}. 
We conducted experiments using both the 
uni-directional and bi-directional Long Short-Term Memories (LSTMs)
\cite{S_Hochreiter_neural_computation_1997_00} in the encoder.
However, uni-directional LSTMs are used in the decoder.
For online speech recognition experiments, we used the
MoChA models \cite{c_chiu_iclr_2018_00} with a chunk size of 2.
 For better stability in LSTM training, we use the gradient clipping by  
global norm \cite{r_pascanu_icml_2013}, which is implemented as  
{\tt tf.clip\_by\_global\_norm } API in Tensorflow  \cite{m_abadi_usenix_2016}.
We use six layers of encoders and one layer of decoder followed by a softmax
layer.

%
%
%
\section{Experimental Results}
\label{sec:experimental_results}
%
In this section, we present a summary of experimental results obtained
for our E2E speech recognition systems built using the proposed
Samsung Research end-to-end training framework.

\begin{table}[!tbhp]
  \renewcommand{\arraystretch}{1.3}
  \centering
        \caption{\label{tbl:power_law_result}
        Word Error Rates (WERs) obtained using MFCC implemented in
        \cite{b_mcfee_proc_scipy_2015_00} and power mel filterbank coefficients
        \\ on the Librispeech corpus \cite{v_panayotov_icassp_2015_00}. For
        each WER number, \\the same experiment was conducted twice and averaged.
        }
        %\vspace{-3mm}
        \begin{tabular}{| c | c  || c | c | c | c |}
          \hline
          \multicolumn{2}{| l ||}{Cell Size}
                                 & \specialcell{MFCC}
                                 & \specialcell{Power Mel \\ 
                                 Filterbank \\
                                 Coefficients} \\
          \hline
          \multirow{3}{*}{1536 cell}    
                  & test-clean  &   4.06  \% &  3.94 \%   \\
                  & test-other  &  13.97  \% & 13.56 \%   \\
                  & average     &   9.02  \% &  8.75 \%   \\
          \hline
       \end{tabular}
       \vspace{-2mm}
\end{table}
%
%
\begin{table}[!tbhp]
  \renewcommand{\arraystretch}{1.3}
  \centering
        \caption{\label{tbl:vtlp_result}
        Word Error Rates (WERs) obtained with VTLP processing
        with different warping factor $\alpha$ distribution, 
        and with and without an RNN LM. The warping factor $\alpha$
			  is the constant controlling warping in \eqref{eq:bilinear_transformation}.
        }
        %\vspace{-3mm}
        \begin{tabular}{| c | c  || c | c | c | c |}
          \hline
          \multicolumn{2}{| l ||}{Warping Factor}
                                 & \specialcell{ 0.7 $\sim$ 1.3 }
                                 & \specialcell{ 0.8 $\sim$ 1.2 }  
                                 & \specialcell{ 0.9 $\sim$ 1.1 }   \\
          \hline \hline
          \multirow{3}{*}{\specialcell{Without \\ RNN-LM}}
                  & test-clean  &   3.82  \% &  3.66  \%   &  3.86  \%  \\  
                  & test-other  &  12.50  \% & 12.39  \%   & 12.35  \%  \\  
                  & average     &   8.16  \% &  8.03  \%   &  8.11  \%  \\  
          \hline
          \multirow{3}{*}{\specialcell{With \\ RNN-LM}}
                  & test-clean  &   2.93  \% &  2.85  \%   &  2.96  \%  \\  
                  & test-other  &  10.40  \% & 10.25  \%   & 10.13  \%  \\  
                  & average     &   6.67  \% &  6.55  \%   &  6.55  \%  \\  
          \hline
       \end{tabular}
       \vspace{-2mm}
\end{table}
%
%
%
For near-field speech recognition experiments, we use the open source 
Librispeech database \cite{v_panayotov_icassp_2015_00}, as well as our in house
Bixby Usage training and test corpuses for English. 
The LibriSpeech dataset consists of around 960 hours of training data
consisting of 281,241 utterances. The evaluation set consists of the 
official 5.4 hours {\tt test-clean} and 5.1 hours {\tt test-other} data.
The Bixby Usage train corpus consists of approximately 10,000 hours of
anonymized Bixby usage data. The evaluation set consists of around 1000 
open domain utterances.

In Table \ref{tbl:power_law_result}, we compare the performance between
the baseline MFCC and the power-law of $(\cdot)^{\frac{1}{15}}$ features 
for a bidirectional full attention (BFA) E2E model with an encoder LSTM 
cell size of 1536. Especially for {\tt test-other}, which
is a more difficult task, the power mel filterbank coefficients 
shows better performance than the baseline MFCC.

In Table \ref{tbl:vtlp_result}, we show Word Error Rates (WERs) for a BFA model using 
different window sizes and warping coefficient distributions, with and without
using an external Recurrent Neural Network (RNN) Language Model (LM) 
\cite{a_zeyer_interspeech_2018_00} built using the standard LibriSpeech LM corpus.
The best performance was achieved when the window length is 50 {\it ms} and 
the warping coefficients are uniformly distributed between 0.8 and 1.2.
We obtained 3.66 \% WER on the {\it test-clean} database and 12.39 \%
WER on the {\it test-other} database without using an LM. 
To best of our knowledge, this is the
best result on this database without using any language models.
Using this shallow-fusion technique 
  with an RNN-LM, we achieved WERs of 2.85 \% and 10.25 \% 
  on the Librispeech {\tt test-clean} and {\tt test-other} databases, respectively.
To the best of our knowledge, the 2.85 \% WER on the {\tt test-clean} 
is the best result ever reported on this database using any approach, 
and 10.25 \% WER on the {\tt test-other} is the best result on this database using 
  an end-to-end recognition system.

In Table \ref{tbl:near_field_result}, we summarize our WER results for 
both the LibriSpeech and Bixby near-field E2E ASR models
with and without using an external RNN-LM trained using around 65GB of 
Bixby LM corpus with an architecture exactly similar to the
LibriSpeech LM model used in \cite{a_zeyer_interspeech_2018_00}.
All the models mentioned in the table have 1024 LSTM cells
in each encoder layer.
For comparison, the best WFST based conventional LSTM-HMM based ASR system
gives a WER of 8.85\% on the Bixby same open domain test set.
We can see that our current Bixby E2E BFA model is $\sim$10\% better, while
our MoChA streaming model is $\sim$10\% poorer compared to
the conventional WFST based DNN-HMM system.
%
\begin{table}[!tbhp]
  \renewcommand{\arraystretch}{1.3}
  \centering
        \caption{\label{tbl:near_field_result}
        Summary of Word Error Rates (WERs) obtained for different
        LibriSpeech and Bixby near-field E2E ASR models 
        with and without an RNN LM.
        }
        %\vspace{-3mm}
        \begin{tabular}{| c | c  || c | c |}
          \hline
          \multicolumn{2}{| l ||}{Models}
                                 & BFA
                                 & MOCHA \\
          \hline \hline
          \multirow{2}{*}{\specialcell{LibriSpeech}}
                  & w/o LM  &   3.66  \% &  6.78  \%  \\  
                  & RNN-LM  &   2.85  \% &  5.54  \%  \\  
          \hline
          \multirow{2}{*}{\specialcell{Bixby}}
                  & w/o LM  &   8.25  \% &  10.77  \% \\  
                  & RNN-LM  &   7.92  \% &   9.95  \% \\  
          \hline
       \end{tabular}
       \vspace{-2mm}
\end{table}
%

The performance our far-field E2E ASR model trained using the proposed
data pipeline with example servers and acoustic simulator is shown 
in Fig.~\ref{fig:plot_farfield_wer}. The performance of the far-field
models are evaluated on a English Commands test set with 900 utterances
and recorded at 4 different distances: 0.5m, 1m, 3m and 5m.
%
\begin{figure}[t]
            \begin{center}
                    {\includegraphics[width=80mm]{../figures/plot_farfield_wer}}
            \end{center}
  %\vspace{-6mm}
 \caption{\label{fig:plot_farfield_wer}
  Speech recognition accuracy with respect to the distance between 
  the microphone and the speaker.
 }
\end{figure}
%

%%
%
%
\section{Conclusions}
\label{sec:conclusions}
We presented a new end-to-end training framework and
 strategies for training 
state-of-the-art end-to-end speech recognition systems.
Our training system utilizes a cluster of Central Processing Units 
  (CPUs) and Graphics Processing Units (GPUs).
The entire data reading, large scale data augmentation,
neural network parameter updates are performed on-the-fly
using example servers and sharded {\tt TFRecords} and {\tt tf.data}. 
  We use vocal tract length perturbation and an acoustic simulator 
for data augmentation. Horovod {\tt allreduce} approach is employed 
to train the neural network parameters using Adam optimizer.
We evaluated the effectiveness of our system on the standard 
Librispeech corpus \cite{v_panayotov_icassp_2015_00} and 
  10,000-hr anonymized Bixby English training 
 and test sets both in near-field as well as
far-field scenarios.  
%\clearpage
%\newpage
\eightpt
\bibliographystyle{IEEEtran}
\bibliography{../../common_bib_file/common_bib_file}

\end{document}