### 2019-09-02 cv_chanwcom/cv_chanwcom.tex
Updates the CV

% generated with the docstrip utility.
%
% The original source files were:
%
% currvita.dtx  (with options: `testfile')
%
% LaTeX sample file for the currvita package
%
% Copyright (C) 1999 Axel Reichert
% See the files README and COPYING.
%
% \CharacterTable
%   {Upper-case    \A\B\C\D\E\F\G\H\I\J\K\L\M\N\O\P\Q\R\S\T\U\V\W\X\Y\Z
%   Lower-case    \a\b\c\d\e\f\g\h\i\j\k\l\m\n\o\p\q\r\s\t\u\v\w\x\y\z
%   Digits        \0\1\2\3\4\5\6\7\8\9
%   Exclamation   \!     Double quote  \"     Hash (number) \#
%   Dollar        \$     Percent       \%     Ampersand     \&
%   Acute accent  \'     Left paren    \(     Right paren   \)
%   Asterisk      \*     Plus          \+     Comma         \,
%   Minus         \-     Point         \.     Solidus       \/
%   Colon         \:     Semicolon     \;     Less than     \<
%   Equals        \=     Greater than  \>     Question mark \?
%   Commercial at \@     Left bracket  \[     Backslash     \\
%   Right bracket \]     Circumflex    \^     Underscore    \_
%   Grave accent  \`     Left brace    \{     Vertical bar  \|
%   Right brace   \}     Tilde         \~}
\documentclass[10pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% packages
\usepackage{color}
\usepackage{currvita}
\usepackage{helvet} % Default font is the helvetica postscript font
\usepackage{hyperref}
\usepackage{mdwlist}
\usepackage[resetlabels]{multibib}
\usepackage{titlesec}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[margin=1.2in]{geometry}

\definecolor{red}{RGB}{128,0,0}
\definecolor{dark_blue}{RGB}{0,0,128}
\hypersetup{
  colorlinks=true,
  linkcolor=dark_blue,
  filecolor=dark_blue,
  urlcolor=dark_blue,
}
\newcommand{\chanwcom}{\color{dark_blue} \underline{\textbf{C. Kim}}}
\newcommand{\lineunder}{
    \vspace*{-15pt}\par\hspace*{-6pt}\color{dark_blue}\hrulefill\\\color{black}}
%\newcites{pubs,intjourn,bookchap,intconf,domconf,thesis,uspatents,uspatentsapp,koreapatents}
\newcites{intjourn,bookchap,intconf,domconf,thesis,uspatents,uspatentsapp,koreapatents}
    {
     %{Peer-reviewed International Journal/Conference Papers \\
     %     \lineunder \vspace{-5mm}},
     {International Jounral Papers \\
       \lineunder \vspace{-5mm}},
     {Book Chapter \\
       \lineunder \vspace{-5mm}},
     {Peer-reviewed International Conference Papers \\
       \lineunder \vspace{-5mm}},
     {Domestic Conference Papers \\ \lineunder \vspace{-5mm}},
     {Theses \\ \lineunder \vspace{-5mm}},
     {US Patents Issued \\ \lineunder \vspace{-5mm}},
     {US Patents Application \\ \lineunder \vspace{-5mm}},
     {Korea Patents Issued \\ \lineunder \vspace{-5mm}}}
\titleformat{\section}[block]{\color{dark_blue}\Large\bfseries\filcenter}{}{1em}{}
\titleformat{\subsection}[hang]{\bfseries}{}{1em}{}

\newenvironment{packed_itemize}{
  \begin{itemize}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
  }{\end{itemize}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand*{\ac}[1]{\mbox{#1}}
\tolerance=600
\begin{document}
\begin{flushright}
  {\footnotesize \it    Curriculum Vitae-Chanwoo Kim}
\end{flushright}
%
\section*{\filleft \LARGE \it  \href{https://scholar.google.com/citations?user=pJoZXxYAAAAJ&hl=en}{Chanwoo Kim} \filleft}
\lineunder
{  \it
56 Seongchon-Gil, Seoul R \& D campus, Seocho-gu, Seoul, 06765, South Korea \\
Cell: +82-10-4043-3996, E-mail:
\href{mailto:chanw.com@samsung.com}{chanw.com@samsung.com} \\ \\
Profile Page at Google Research : \href{https://research.google.com/pubs/105202.html}
  {\url{https://research.google.com/pubs/105202.html}} \\ \\
Google Scholar page : \href{https://scholar.google.com/citations?user=pJoZXxYAAAAJ&hl=en}
  {Link to the page} {\color{red} (h-index :22, citations: 1616)} \\ \\
Personal homepage: \href{
  http://www.cs.cmu.edu/~chanwook}{\url{http://www.cs.cmu.edu/~chanwook}}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Education
\section*{Education}
\lineunder
{\sl  Ph. D.} \\
{\color{dark_blue} \href{http://lti.cs.cmu.edu}{Language Technologies Institute},
\href{http://cs.cmu.edu}{School of Computer Science}\\
Carnegie Mellon University}, Pittsburgh, PA \hfill Sept. 2005 - Dec. 2010 \\
Advisor: Prof. Richard Stern \\
THESIS - \href{http://www.lti.cs.cmu.edu/research/thesis/2011/chanwoo_kim.pdf}{Signal Processing for Robust Speech Recognition Motivated by
Auditory Processing} \\ \\
%
{\sl  M. S.} \\
{\color{dark_blue} School of Electrical Engineering and Computer Science} \\
\href{http://www.snu.ac.kr}{Seoul National University}, Seoul, Korea \hfill March 1998 - Feb. 2001 \\
Advisor: Prof. Wonyong Sung \\
THESIS - \href{http://library.snu.ac.kr/search/DetailView.ax?sid=6&cid=925773}{Implementation of an Intonation and Pronunciation Checking System for Embedded Systems} \\ \\
%
{\sl  B. S.} \\
{\color{dark_blue} School of Electrical Engineering} \\
\href{http://www.snu.ac.kr}{Seoul National University}, 
Seoul, Korea \hfill March 1994 - Feb. 1998 \\
Thesis advisor: Prof. Min-Koo Han \\ \\
%
{\sl  High School} \\
\href{http://www.sshs.hs.kr/}{Seoul Science High School}, Seoul, Korea \hfill March 1991 - Feb. 1994 \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Research Interests
\section*{Research Interests}
\lineunder
\begin{packed_itemize}
\item Automatic speech recognition
\item Machine Learning and deep learning for signal processing
\item Training deep-learning system using simulated data
\item Speech enhancement (using a single microphone, dual microphones, and multiple microphones)
\item Human auditory processing
\end{packed_itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Award
\section*{Award}
\lineunder
\begin{packed_itemize}
\item {\color {dark_blue} Google spot bonus},
    Acoustic modeling for Google Home. Dec. 2016.
\item {\color {dark_blue} Google peer bonus},
    Member of the summit squad for the Google Pittsburgh Ph. D. summit 2014, Google, Jan. 2014.
\item {\color{dark_blue} Google peer bonus},
    Design and implementation of the room simulator and robustness training and evaluation using it, Google, Dec, 2013
\item {\color {dark_blue} 17-th Samsung Humantech Thesis Bronze prize},
    {\color{dark_blue} C. Kim} and R. M. Stern,
    "Power-Normalized cepstral coefficients for robust speech recognition",
     Feb. 2011
\item {\color {dark_blue} 16-th Samsung Hamantech Thesis Honour prize},
    {\color{dark_blue} C. Kim} and R. M. Stern,
    "Small power boosting and spectral subtraction for robust speech
     recognition",  Feb. 2010
\item {\color{dark_blue} Research fellowship for Ph. D. study} 2005 - 2010, Carnegie Mellon University, Jan. 2005.
\item {\color{dark_blue} Scholarship for studying abroad 2005 - 2006},
    Korea Institute of Energy Evaluation and Planning, July, 2005.
\item {\color{dark_blue} Graduate cum laude} with a bachelor’s degree, Seoul National University, Feb. 1998
\end{packed_itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Employment
\section*{Employment}
\lineunder
%
{\color{dark_blue} Samsung Research} Speech Processing Lab. \hfill Feb. 2018
- Current \\
56 Seongchon-Gil Seoul R \& D Campus, Seoul, 06765, Korea \\
Corporate Vice President \\
\vspace{-5mm}
\begin{itemize*}
  \item
    {\color{dark_blue} On-device and server neural end-to-end speech recognition} \\
  \item
    {\color{dark_blue} Wake-up engine} \\
  \item
    {\color{dark_blue} Far-field processing} \\
  \item
    {\color{dark_blue} Neural Text-to-speech engine development} \\
\end{itemize*}
%
{\color{dark_blue} Google} Google Speech \hfill Feb. 2013 - Feb. 2018 \\
1600 Amphitheatre Pkwy Mountain View, CA, 94043  \\
Senior Software Engineer \\
\vspace{-5mm}
\begin{itemize*}
  \item
    {\color{dark_blue} Room simulation system for acoustic model simulation} \\
      Designed and implemented the room acoustics simulation system using
      the image method for various kinds of real room environments. Using
      deep learning techniques, acoustic models were trained using data generated by this
      room acoustics simulator. This system has reduced Word Error Rate to half [C. Kim. et. al., Interspeech 2017]. This system is used for commercialized Google
      speech recognition systems. \\

  \item
    {\color{dark_blue} Developing next-generation voice-assistant devices for Google}
     Have participated and lead efforts for the future voice assistant
     devices at Google. Currently under development. \\

  \item
    {\color{dark_blue} Robust speech recognition for Google Home} \\
      Have participated in the Google Home project by generating
      simulated training sets and performing training and evaluation.
      This system was launched and commercialized. \\

  \item
    {\color{dark_blue} Noise robustness for speech recognition using deep learning} \\
      Developed a system which learns the mapping from noisy feature to clean feature using
      deep learning techniques for feature enhancement and noise robustness
      of speech recognition systems.  \\

  \item
    {\color{dark_blue} Multi-style training for speech recognition systems} \\
      Trained acoustic models for speech recognition systems using various
      machine learning techniques and the aforementioned room simulator.
      Using this Multi-style TRaining (MTR) approach, we obtained remarkable
      improvement under noisy environments. \\

  \item
    {\color{dark_blue} Development of the PAINT algorithm for robust speech recognition} \\
      Developed a sound source separation algorithm using
      multi-microphones named the Power Angle Information Near
      Target (PAINT) algorithm for future Internet Of Things
      (IOT) products \\
  \item
    {\color{dark_blue} Development of the TMT algorithm for robust speech recognition} \\
      Developed the Temporal Masking and Threshold (TMT)algorithm motivated by the on-set enhancement
      techniques for robust speech recognition under reverberant environment. \\

  \item
    {\color{dark_blue} Development of a Signal-to-Noise Ratio (SNR) estimation algorithm} \\
      Developed a modified Waveform Amplitude Distribution Analysis (WADA)
      algorithm as a Mapper in MapReduce at Google. It has been a valuable
      tool at Google speech.

\end{itemize*}
\vspace{5mm}
%
{\color{dark_blue} Microsoft} Microsoft Speech  \hfill Jan. 2011 - Jan. 2013\\
One Microsoft Way, Redmond, WA, 98052  \\
Speech Scientist \\
\vspace{-5mm}
\begin{itemize*}
  \item
    {\color{dark_blue} Development of the SMAD algorithm} \\
      Developed a sound source separation algorithm named Statistical
      Modeling of Angle Distribution (SMAD) for multi-microphone
      sound source separation \\
  \item
  {\color{dark_blue} Distributed processing for speech recognition} \\
      Participating in developing the infrastructure for processing big data
      for speech recognition using Cosmos. \\
  \item
    {\color{dark_blue} Post-processing of Gaussian Mixture Model (GMM)} \\
      Introducing an approach to balance deletion/insertion errors
      by modifying the mixture components of GMM. \\
  \item
    {\color{dark_blue} {Audio-Video (AV) capture pipeline for Microsoft Windows Phone}} \\
      Audio/Video capturing using Microsoft media foundation for Windows Phone 8. \\

  \item
    {\color{dark_blue} Media Foundation Transforms (MFTs)}
        Design and implementation of several Media Foundation Transforms
      (SW Jpeg Encoder MFT, Demux MFT, etc). \\
  \item
    {\color{dark_blue} Camera pipeline} \\
      Working for EXIF processing, meta data handling etc.
\end{itemize*}
\vspace{5mm}
%
{\color{dark_blue} LG Electronics } \hfill June 2003 - Aug. 2005 \\
128, Yeoui-daero Yeongdeungpo-gu, Seoul, Korea, 150-721  \\
Senior Research Engineer, VoIP Phone Conference Call Development Team Leader \\
\vspace{-5mm}
\begin{itemize*}
  \item {\color{dark_blue} VoIP handset project}
      Manging the conference-call part using the Session Initiation Protocol (SIP). \\

  \item {\color{dark_blue} Video On Demand (VOD) system project}
      Participating in the VOD project working on streaming protocols such as
      Real-Time Protocol (RTP), Real-Time Control Protocol (RTCP), and
      Real-Time Streaming Protocol (RTSP). \\

  \item {\color{dark_blue} Development of an audio/video synchronization algorithm}
      Developing an audio/video synchronization algorithm for
      Video Telephony (VT) using Network Time Protocol (NTP). \\

  \item {\color{dark_blue} Testing and evaluation of multimedia codes}
      H.263, H.264, AAC, AAC+, and H.323 \\

  \item {\color{dark_blue} Multimedia platform survey}
      Texas Instruments (TI) Open Multimedia Application Platform (OMAP), Neomagic MiMagic6, Renesas SH-Mobile V. etc
\end{itemize*}
\vspace{5mm}
%
{\color{dark_blue} SK Teletech } \hfill June 2003 - Aug. 2005 \\
Research Engineer \\
\\
{\color{dark_blue} Edumediatek } \hfill June 2000 - July 2002 \\
Research Engineer \\
A research venture company established by Prof. Wonyong Sung at Seoul National University
\begin{itemize*}
  \item {\color{dark_blue} Computer Aided Language Learning (CALL) System} Intonation and vowel checking
system for embedded systems \\
  \item {\color{dark_blue} Development of a robust formant extraction algorithm} Combination of the spectral \\
 peak picking and root solving \\
  \item {\color{dark_blue} Development of a robust pitch extraction algorithm}
\end{itemize*}
\vspace{5mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Graduate research
\section*{Graduate research}
\lineunder

\begin{itemize*}
  \item {\color{dark_blue} Power Normalized Central Coefficients (PNCC)}
      Motivated by human auditory processing, this new feature set
      incorporates modulation frequency, temporal masking, and
      rate-nonlinearity concepts. The features require low computation
      and on-line implementation is possible. \\

  \item {\color{dark_blue} Small Power Boostin (SPB) Algorithm}
      Developed the SPB algorithm motivated by the human rate-intensity
      as well as temporal and spectral smoothing. This algorithm works
      especially well for non-stationary noise such as music noise. \\

  \item {\color{dark_blue} Dual-microphone speech enhancement systems}
      Developed the Phase Difference Channel Weighting (PDCW) algorithm
      which performs sound source separation without a priori knowledge
      of source locations. \\

  \item {\color{dark_blue} Single-microphone speech enhancement system}:
      Developed the Power-function-based Power Distribution Normalization
      (PPDN) algorithm which enhances speech. \\

  \item {\color{dark_blue} Automatic speech recognition with
      low computational cost}
      Focused on developing online algorithms, which are robust against
      noise with low computational complexity. \\

  \item {\color{dark_blue} SNR Estimation algorithm for speech}:
      Developed a new algorithm which estimates Signal-to-
      Noise Ratio (SNR) more accurately than the standard NIST STNR algorithm. \\

  \item {\color{dark_blue} Automatic ITD threshold detection algorithm}:
      Developed a new algorithm which can obtain the optimal ITD threshold
      for spatial separation of signals. \\

  \item {\color{dark_blue} Dereverberation algorithm}
      Developed Suppression of Slowly-varying and Falling edges (SSF): a
      simple algorithm for robust speech recognition that is highly effective
      in reverberant environments. \\

  \item {\color{dark_blue} HLab C++ automatic speech recognition system}
      Developed a C++ HMM (Hidden Markov Model) speech recognition system from
      scratch. \\

  \item {\color{dark_blue} Power Normalized Voice Activity Detection (PN-VAD)}
      Developed a new voice activity detection algorithm that is very accurate
      and computationally efficient. \\

  \item {\color{dark_blue} DARPA GALE project}
      Robust front end for the DARPA GALE project. Applied a new feature set
      that we developed for the 180-hr training set and the large vocabulary
      Broadcast News test set. \\

  \item {\color{dark_blue} NSF auditory signal processing project}
      Developed features motivated by auditory processing, binaural processing,
      and poly-aural processing. \\

  \item {\color{dark_blue}Samsung speech-to-speech translator project}
      Developed low-complexity online algorithms for embedded processors.
      Developed an on-line noise-robustness feature extraction algorithm,
      a binaural source-separation algorithm, an online VAD, an online MVN,
      and supporting technologies.
      Applied the algorithm to a 64,000-word database with a 500-hour training set. \\

  \item {\color{dark_blue} Voice activity detector project}
      Designed a robust VAD which requires very low computation.

\end{itemize*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Invited talk
\section*{Invited Talks}
\lineunder
\begin{itemize*}
  \item ``Introduction to contemporary speech recognition systems and recent advances from deep-learning and human auditory processing techniques," Korea Aerospace University, Goyang, Korea, May. 2015.
  \item ``Research direction for robust speech recognition," Samsung Advanced Institute of Technology, Suwon, Korea, Dec. 2014.
  \item ``Speech recognition research using neural networks," Seoul National University, Seoul Korea, Sept. 2014.
  \item ``Speech Noise Robustness for Intelligent Cars and Home Appliances," Yonsei University, Wonju, Korea, Sept. 2014.
  \item ``TMT, SSF, and PAINT algorithms for far-field speech recognition," Sogang University, July, 2014.
  \item ``Speech Recognition Infrastructure and Noise Robustness Efforts," Seoul National University. July, 2014.
  \item ``Signal Processing for Robust Speech Recognition Motivated by Auditory Processing," Sogang University, Sept. 2012.
\end{itemize*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Invited talk
\section*{Teaching Experience}
\lineunder
\begin{itemize*}
  \item Teaching Assistant, 18-792 Advanced Digital Signal Processing,
    Electrical and Computer Engineering Department, Carnegie Mellon University, Spring, 2010.
  \item Teaching Assistant, including weekly recitations,
    15-211 fundamental data structures and algorithms,
    School of Computer Science, Carnegie Mellon University, Fall, 2009.
\end{itemize*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Publications
% 2019
\nociteintconf{d_gowda_interspeech_2019_00}
\nociteintconf{c_kim_interspeech_2019_00}
\nociteintconf{a_menon_icassp_2019_00}

% 2018 
\nociteintconf{c_kim_interspeech_2018_00}
\nociteintconf{x_zhou_isnn_2018_00},
\nociteintconf{c_kim_icassp_2018_00}
\nociteintconf{c_kim_icassp_2018_01}
%\nociteintconf{r_nongpiur_icassp_2018_00}
%\nociteintconf{a_menon_icassp_2018_00}
% 2017
\nociteintconf{a_menon_asru_2017_00}
\nociteintconf{C_Kim_INTERSPEECH_2017_1}
\nociteintconf{B_Li_INTERSPEECH_2017_1}
\nociteintconf{A_Menon_INTERSPEECH_2017_1}
\nociteintjourn{T_Sainath_IEEETran_2017_1}
\nocitebookchap{T_Sainath_Book_Chapter_2017_1}
% 2016
\nociteintjourn{C_Kim_IEEETran_2016_1}
\nociteintjourn{B_Cho_SigProcLetters_2016_1}

% 2015
\nociteintconf{C_Kim_INTERSPEECH_2015}

% 2014
\nociteintconf{C_Kim_INTERSPEECH_2014_2}
\nociteintconf{C_Kim_INTERSPEECH_2014_1}

% 2013

% 2012
\nociteintconf{C_Kim_ICASSP_2012_2}
\nociteintconf{C_Kim_ICASSP_2012_1}

% 2011
\nociteintconf{C_Kim_ICASSP_2011_3}
\nociteintconf{C_Kim_ICASSP_2011_2}

% 2010
\nociteintconf{C_Kim_ICASSP_2010_1}
\nociteintconf{C_Kim_INTERSPEECH_2010_2}
\nociteintconf{C_Kim_INTERSPEECH_2010_1}

% 2009
\nociteintconf{C_Kim_ASRU_2009_2}
\nociteintconf{C_Kim_ASRU_2009_1}
\nociteintconf{C_Kim_INTERSPEECH_2009_2}
\nociteintconf{C_Kim_INTERSPEECH_2009_1}

% 2008
\nociteintconf{C_Kim_INTERSPEECH_2008}
\nociteintconf{R_M_Stern_HSCMA_2008}

% 2007

% 2006
\nociteintjourn{C_Kim_EurasipJourn_2006}
\nociteintjourn{C_Kim_IEICE_2006}
\nociteintconf{C_Kim_INTERSPEECH_2006}
\nociteintconf{C_Kim_ICCE_2006}

% 2005
\nociteintjourn{C_Kim_TransCE_2005}
\nociteintconf{C_Kim_ICCE_2005}

% 2004

% 2003

% 2002
\nociteintconf{C_Kim_INTERSPEECH_2002}

% 2001
\nociteintconf{C_Kim_ICSP_2001}

\bibliographystyleintjourn{ieee_tran}
\bibliographyintjourn{../common_bib_file/common_bib_file}

\bibliographystylebookchap{ieee_tran}
\bibliographybookchap{../common_bib_file/common_bib_file}

\bibliographystyleintconf{ieee_tran}
\bibliographyintconf{../common_bib_file/common_bib_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Domestic conference
\nocitedomconf{C_Kim_KCC_2005}

\bibliographystyledomconf{ieee_tran}
\bibliographydomconf{../common_bib_file/common_bib_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Theses
\nocitethesis{C_Kim_PhDThesis_2010}
\nocitethesis{C_Kim_MSThesis_2001}

\bibliographystylethesis{ieee_tran}
\bibliographythesis{../common_bib_file/common_bib_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% US Patents Issued
% 2016
\nociteuspatents{C_Khwand_USPatent_Issued_2016_1}

% 2015
\nociteuspatents{C_Kim_USPatent_Issued_2015_1}

% 2014
\nociteuspatents{C_Kim_USPatent_Issued_2014_1}

% 2011
\nociteuspatents{C_Kim_USPatent_Issued_2011_1}

% 2010
\nociteuspatents{C_Kim_USPatent_Issued_2010_2}
\nociteuspatents{C_Kim_USPatent_Issued_2010_1}

% 2009
\nociteuspatents{C_Kim_USPatent_Issued_2009_4}
\nociteuspatents{C_Kim_USPatent_Issued_2009_3}
\nociteuspatents{C_Kim_USPatent_Issued_2009_2}
\nociteuspatents{C_Kim_USPatent_Issued_2009_1}

% 2008
\nociteuspatents{C_Kim_USPatent_Issued_2008_1}

\bibliographystyleuspatents{ieee_tran}
\bibliographyuspatents{../common_bib_file/common_bib_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% US Patents Application

% 2016
\nociteuspatentsapp{A_Narayanan_USPatent_Application_2016_1}
\nociteuspatentsapp{A_Misra_USPatent_Application_2016_1}

% 2014
%\nociteuspatentsapp{C_Kim_USPatent_Application_2014_1}
%\nociteuspatentsapp{C_Khawand_USPatent_Application_2014_1}

% 2012
\nociteuspatentsapp{C_Kim_USPatent_Application_2012_1}

% 2006
\nociteuspatentsapp{C_Kim_USPatent_Application_2006_5}
\nociteuspatentsapp{C_Kim_USPatent_Application_2006_4}
\nociteuspatentsapp{C_Kim_USPatent_Application_2006_3}
\nociteuspatentsapp{C_Kim_USPatent_Application_2006_1}

% 2005
\nociteuspatentsapp{C_Kim_USPatent_Application_2005_3}
\nociteuspatentsapp{C_Kim_USPatent_Application_2005_1}

\bibliographystyleuspatentsapp{ieee_tran}
\bibliographyuspatentsapp{../common_bib_file/common_bib_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Korea Patents Issued

% 2007
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2007_1}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2007_2}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2007_3}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2007_4}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2007_5}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2007_6}

% 2006
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2006_1}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2006_2}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2006_3}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2006_4}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2006_5}
% The following has become void.
%\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2006_6}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2006_7}
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2006_8}

% 2005
\nocitekoreapatents{C_Kim_KoreaPatent_Issued_2005_1}

\bibliographystylekoreapatents{ieee_tran}
\bibliographykoreapatents{../common_bib_file/common_bib_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Technical skills
\section*{Technical skills}
\lineunder
\begin{itemize*}
  \item Programming Language :
      C, C++, Pyhon, Perl, Java, Matlab, various shell scripts, etc.
  \item Programming Environments :
      Win32 API, Microsoft Foundation Class (MFC), NumPy/SciPy, MapReduce, and various Linux environments.
  \item Speech Recognition Engine :
      Sphinx, Hidden markov model Tool Kit (HTK), Janus, etc.
  \item Development Tool : Various Linux tools including gcc. Microsoft visual studio,
      Arm Developer Suite (ADS), Code Composer Studio (CCS), etc.
  \item Real-time OS : Rex, Nucleus
\end{itemize*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% References
\section*{References}
\lineunder

\begin{itemize*}
  \item
    {\color{dark_blue} Prof. Richard M. Stern} (Ph. D. study advisor 2005-2010)\\
    Department of Electrical and Computer Engineering,
    Carnegie Mellon University, \\
    B24 Porter Hall \\
    5000 Forbes Ave, Pittsburgh, PA 15213, USA (Tel: +1-412-268-2535) \\
    E-mail: \href{mailto:rms@cs.cmu.edu}{rms@cs.cmu.edu} \\

  \item
    {\color{dark_blue} Prof. Wonyong Sung} (M. S. study advisor 1999-2001) \\
    Department of Electrical and Computer Engineering, Seoul National Univ. \\
    1 Gwanangno, Gwanak-gu, Seoul, Korea, 151-744 (Tel: +82-2-880-1816)  \\
    E-mail: \href{mailto:wysung@dsp.snu.ac.kr}{wysung@dsp.snu.ac.kr} \\

  \item
    {\color{dark_blue} Prof. Bhiksha Raj} (Collaborating professor) \\
    Language Technologies Institute, School of Computer Science,
    Carnegie Mellon University, \\
    6705 Gates-Hillman Center,
    5000 Forbes Ave, Pittsburgh, PA 15213  (Tel: +1-412-268-9826) \\
    E-mail: \href{mailto:bhiksha@cs.cmu.edu}{bhiksha@cs.cmu.edu} \\

  \item {\color{dark_blue} Prof. Rita Singh} \\
    Language Technologies Institute, School of Computer Science,
    Carnegie Mellon University, \\
    6723 Gates-Hillman Center,
    5000 Forbes Ave, Pittsburgh, PA 15213  (Tel: +1-412-268-9859) \\
    E-mail: \href{mailto:rsingh@cs.cmu.edu}{rsingh@cs.cmu.edu} \\

  \item {\color{dark_blue} Prof. Kwang-deok Seo} (Manager at LG 2003-2005) \\
    Computer\&Telecommunications Engineering Division,
    Yonsei University, \\
    1 Yonseidae-gil, Wonju, Gangwon-do, 220-710, Korea, (Tel: +1-33-760-2788) \\
    E-mail: \href{mailto:kdseo@yonsei.ac.kr}{kdseo@yonsei.ac.kr}
\end{itemize*}
\vspace{10mm}
\begin{flushright}
    Updated Sept. 1st , 2017
\end{flushright}
\end{document}
\endinput
%%
%% End of file `cvtest.tex'.