### 2019-09-14 nips_2018_neural_front_end/doc/main_bak.tex

%% bare_jrnl.tex
%% V1.4
%% 2102/12/27
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/10 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex,
%%                    bare_jrnl_transmag.tex
%%*************************************************************************

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEalign family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{dsfont}
\usepackage{tikz}
\usetikzlibrary{arrows,decorations.markings}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\large]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% New commands
\newcommand{\chanwcom}{{C. Kim}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Phase Difference Smoothing for Sound Source Separation Under Reverberant Environment}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{
        %Chanwoo~Kim,~\IEEEmembership{Member,~IEEE,}
        %and~M~Bacciani,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
\thanks{M. Shell is with the Department
of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised December 27, 2102.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~11, No.~4, December~2102}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2102 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This paper presents a new algorithm for normalizing multi-microphone
signals under reverberation. In this study, we observe that in
reverberant environments, Interaural Time Delay (ITD) or Phase
Difference (PD) estimation may be distorted due to reflections. We
propose a method referred to as Multi-Microphone Signals Normalization (MMSN) 
to model this distortion as a transfer function 
between multi-microphone signals and to compensate this distortion.
We combine this MMSN algorithm with our previous sound source separation 
algorithm referred to as Phase Difference Channel Weighting (PDCW). 
Speech recognition experiment shows that MMSN algorithm significantly 
enhances robustness of sound source separation algorithms under 
reverberant environments.

\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Multi-microphone signal processing, masking, reverberation,
interaural time delay, phase difference, beam forming,
robust speech recognition
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results.}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Introduction}
\cite{J_Allen_JASA_1979}
\IEEEPARstart{S}{peech} recognition systems have significantly improved
these days, and we have been widespread usage of them
for hand-held devices like smart phones. 
Recent advances in machine learning techniques such 
as Deep Neural Network \cite{G_Hinton_IEEE_Signal_Process_Mag_2012}
and enhanced computational power \cite{V_Vanhoucke_Deep_Learning_NIPS_Workshop_2011}
have made tremendous contribution for improving speech recognition accuracy.
In spite of these success with close-talking applications
 like smart phones, they have been less successful for 
far-field applications such as
speech recognition inside a car or speech recognition for
home appliance devices. The major problem is that under
noisy or reverberant environments, speech recognition
accuracy is quite adversely affected.

Thus far, many researchers have proposed various kinds of algorithms
to address this problem  \cite{P_J_Moreno_ICASSP_1996,
R_M_Stern_ETRW_1997, C_Kim_ASRU_2009_2, C_Kim_ICASSP_2012_1,
C_Kim_ICASSP_2012_2}. To some degrees, these efforts have brought
some improvement. Nevertheless for highly non-statioary noise
or reverberent speech, the improvment has been still limited.
For these environments, recent work motivated by human 
auditory processing seems to be more promising
\cite{C_Kim_ICASSP_2012_1, C_Kim_INTERSPEECH_2010_2, 
C_Kim_ICASSP_2010_1, C_Kim_INTERSPEECH_2006}.

TODO(chanwcom) Revise the introduction
An alternative approach is signal separation based on anal-
ysis of differences in arrival time [8], [9], [10]. It
is well documented that the human binaural system has a
remarkable ability to separate speech and other sounds that
arrive from different azimuths [11]. It has been observed
that various types of cues are used to segregate the target signal
from interfering sources. Motivated by these observations,
many models and algorithms have been developed using
inter-microphone time differences (ITDs), inter-microphone
intensity difference (IIDs), inter-microphone phase differences
(IPDs), and other cues [8], [9], [12], [13]. IPD and ITD
have been extensively used in binaural processing because
this information can be easily obtained by spectral analysis
[12] [14] [15]. ITD can be estimated using either phase
differences [15], cross-correlation [16], [17], or
zero-crossings  [9].

Research on signal separation based on analysis of 
differences in arrival time has be conducted for a long time
by many researchers
\cite{SrinivasanEtAl06, H_Park_SpeechComm_2009, 
P_Aarabi_IEEETranSysManCyber_2005}.

signal separation based on analysis of differences in arrival time 
[1, 2]. It is well known that the human binaural system is remark-
able in its ability to separate speech from interfering sources (e.g.
[3]). Motivated by these observations, many models and algorithms
have been developed using interaural time differences (ITDs), inter-
aural intensity difference (IIDs), interaural phase differences (IPDs),
and other cues [1, 2, 4]. IPD and ITD have been extensively
used in binaural processing because this information can be easily
obtained by spectral analysis [5].

TODO(chanwcom) Mention the frequency smoothing.
PNCC, my thesis, other binaural papers.

In PDS altorithm, 
In MMSN algorithm, to capture the effect of reverberation, we use
a longer window than those used for speech feature extraction.
Based on experimental results, we use a medium-duration window as in
\cite{C_Kim_ASRU_2009_1, C_Kim_ASRU_2009_2, 
C_Kim_PhDThesis_2010, C_Kim_ICASSP_2012_1}.

This paper is organized as follows:

In our previous work TODO[XX] \cite{}, we have frequently assumed that the
each microphone in a microphone array has exactly the same frequency 
response.

TODO
however, this is not true with real-recordings since each microphone
in a multi-microphone array does not have exactly the same

[CWP]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section
\section{Background theory}

\subsection {Near-field effect}

In this work, we assume that we do not have prior knowledge about the
target sound source location. However, based on our

We use ILD, but if we directly use ILDs, then it will

Thus, we try the following feature combinations.

Smoothed

We try the following different features.

log(A / B)
(A / B) ^ (p)

As will be described in TODO(chanwcom)

However, along with this, based on discussion... We assume that
log(A) log(B)
log(A / B)
(A / B) ^ (p)

[CWP]

\subsection {PDCW algorithm}

\subsection {Image method to model reverberation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section
\section{The effects of reverberation on multi-microphone signals}

\subsection{Virtual sources in symmetric and asymmetric target location configuration}

\subsection{The effect of reverberation on the target source}

\begin{figure}[!t]
  \centering
    \resizebox{60mm}{!}{\input{../figures/symmetric_room.tex}}
    \caption {
      A target source and its mirrored image   
    }   
\label{fig:symmetric_configuration}
\end{figure}

\begin{figure}[!t]
  \centering
    \resizebox{60mm}{!}{\input{../figures/asymmetric_room.tex}}
    \caption {
      A target source and its mirrored image   
    }   
\label{fig:asymmetric_configuration}
\end{figure}

\begin{figure}[!t]
  \centering
    \resizebox{60mm}{!}{\input{../figures/one_source_one_wall.tex}}
    \caption {
      A target source and its mirrored image   
    }   
\label{fig:one_target_one_virtual}
\end{figure}

\begin{figure}[!t]
  \centering
    \resizebox{60mm}{!}{\input{../figures/one_source_in_room.tex}}
    \caption {
      A target source and its mirrored image   
    }   
\label{fig:one_target_one_virtual}
\end{figure}

Suppose that we place two microphones inside a room as shown in Fig. XXX.
We refer the two signals received by these microphones as 
$x_0[n]$ and $x_1[n]$, and the target signal as $x[n]$.
Let us further assume that the reflection coefficients of the walls
of the room is a constant. Under this assumption, the effect of 
reflection is frequently modeled by the image method. XXXX

In the image method \cite{J_Allen_JASA_1979}, the RIR from the
source to the receiver is obtained by considering images
sources on an infinite grid of mirrored rooms. In Fig. XX, 
we show a 3 x 3 subset of

In Fig. XXX, we represent a true sound source by a solid circle and
virtual sound sources that are mirrored images of the target source 
by a circle with dashed patterns.

For all sound sources including both the original sound source 
and the virtual sound sources, the impulse response for the
microphone 0 and the microphone 1 may be represented by  
the following equation:

\cite{E_A_Lehmann_ASPAA_2007},

In this case, the signal from the left and right microphones 
$x_l[n]$ and $x_r[n]$ may be represented by the following 
equations:

In Fig. XX, we assume that the mirrored source $\mathcal{M}$
is located along a line with a angle of $\theta_M$ with respect to
the line connecting the target and the center of two microphones. 
In this case, the Intermicrophone Time Difference is given by TODO(chanwcom) [XX][XX]
\begin{align}
  d = \frac{A}{B} 
\end{align}
As a specific example, if the $d_t$ is two meters, and the distance 
then, $d_m$ is

When the reflection coefficient is $r$, after the sound wave hits
the wall, the amplitude of the reflected wave becomes $r$ times 
the original waveform [XX].
The amplitude of the waveform is inversely proportional to 
the distance. Thus, the 
\begin{align}
  \alpha = \frac{d_l}{d_m}  
\end{align}

The ratio of the amplitude from the mirrored source to the target
source $\gamma$ is given by:

\begin{subequations}
  \begin{align}
    x_0[n] & = x_t[n] + r \frac{d_l}{d_m} x_t[n - \tau_0] \\
    x_1[n] & = x_t[n] + r \frac{d_l}{d_m} x_t[n - \tau_1]
  \end{align}
\end{subequations}

Performing STFT, we obtain the following equations:
\begin{subequations}
  \begin{align}
    X_0[m, e^{j \omega_k}] 
    = \left(1  + \gamma e^{-j \omega_k d_0} \right) X_t[m, e^{j \omega_k}] \label{eq:target_left_spec}\\
%
    X_1[m, e^{j \omega_k}] 
    = \left(1  + \gamma e^{-j \omega_k d_1} \right) X_t[m, e^{j \omega_k}] \label{eq:target_right_spec}
  \end{align}
\end{subequations}
where $r$ is the reflection coefficient.

\subsection{The effects of mirrored sources on phase difference}

The relationship between the $T_{60}$ and the reflection coefficient may
be approximated by Eryling's empiricial equation [XX]:
\begin{align}
  \alpha & = 1 - \exp \left(\frac{- 0.16 V}{S T_{60}}  \right) \\
  r & = \sqrt{1 - \alpha^2}
\end{align}
where $\alpha$ and $r$ are an absorption coefficient and a reflection coefficient
respectively, and where $V$ is the volume of the room and $S$ is the total
surface area of the room.  As a specific example, for a typical room of a size 
of 5 m x 4 m x 3 m and and a reverberation time $T_{60}$ of 300 $ms$, the reflection
coefficient $r$ is 0.9575.

When there is only one target without noise, the phase difference between
\eqref{eq:target_left_spec} and \eqref{eq:target_right_spec} is given by:
%
%
\begin{align}
  \Delta \phi[m, e^{j \omega_k}] = \angle \left( 1  + \gamma e^{-j \omega_k \tau_0}\right)
    - \angle \left( 1  + \gamma e^{-j \omega_k \tau_0}\right)
\end{align}
%

Let us define 
\begin{align}
  \tau_{ave} = \frac{\tau_0 + \tau_1}{2}
\end{align}

\begin{align}
  \Delta \phi[m, e^{j \omega_k}] 
        = \tan^{-1}\left( \frac{\gamma \sin \left(\omega_k d_0 \right)}
        {1 + \gamma \cos \left( \omega_k d_0 \right)} \right) \nonumber  \\
        - \tan^{-1}\left( \frac{\gamma \sin \left(\omega_k d_1 \right)}
        {1 + \gamma \cos \left( \omega_k d_1 \right)} \right) 
          \label{eq:phase_diff_one_target_one_mirror}
\end{align}
If the argument $x$ of $\tan^{-1}$ is small enough, then it may be approximated 
by $\tan^{1}(x) \approx x$. In this case, \eqref{eq:phase_diff_one_target_one_mirror},
may be approximated by the following equation:
\begin{align}
  \Delta \phi[m, e^{j \omega_k}] & \approx 
    \gamma \left[ \sin \left(\omega_k d_0 \right) 
    - \sin \left(\omega_k d_1 \right) \right]  \nonumber \\
        & = \gamma \sin \left(\frac{\omega_k \left(d_0 + d_1 \right) }{2}  \right)
    \sin \left( \frac{ \omega_k \left(d_0 - d_1 \right) }{2} \right)
      \label{eq:approx_phase_diff_one_target_one_mirror}
\end{align}

%
From the equation \eqref{eq:approx_phase_diff_one_target_one_mirror}, we observe
that the high frequency component is due to the following term: 
\begin{align}
  \Delta = \sin \left( \frac{\omega_k \left(d_0 + d_1\right)}{2} \right)
\end{align}
With respect to the $\omega$ the period is $x$ given by:
\begin{align}
  \frac{2 \pi}{\tau_{ave}} \label{eq:frequency_domain_period}
\end{align}
From \eqref{eq:frequency_domain_period}, we observe that the period depends on the 
distance between ... . Thus, it is safe to say that 
As distance of the becomes farther, the period becomes smaller.
We assume that $d_{ave}  = \frac{d_0 - d_1}{2}$ is at least 1 meter.

In the real situation of far-field speech recognition,

\subsection{The effects of mirrored sources when the target is inside a room}

where $\omega_k$ is the discrete-frequency index defined by
\begin{align}
  \omega_k = \frac{2 \pi k}{N}, \qquad 0 \le k \le N - 1
\end{align}
where $N$ is the DFT size. Typical reverberation time in terms of
$T_{60}$ is between 100 ms and 400 ms, so when performing the STFT
, the window length should be sufficiently long to capture at least 
the early reflections.

In our previous work XXXX we had assumed that if the target 
sound source is located along the perpendicular bisector to the
line connecting two microphones, the ITD and phase difference
should be zero. However, as may be seen in \eqref{eq:microphone_signals_stft}

\subsection{The effect on the reverberation on the target and the noise source}

In this section, we consider the case where there is a target and 
a noise source inside the room simultaneously.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section
\section{Normalization of multi-microphone signals}

\subsection{Off-line normalization algorithm}

Since $a_{[u, v]}$ and $b_{[u, v]}$ are constants
as long as the target is stationary,  $H_{01}[e^{j\omega_k}]$
is also a constant if the target is stationary. However,
if we calculate $H_{01}[m, e^{j\omega_k}]$ for each frame 
independently for multi-channel signal compensation, 
then as will be seen in XXXX, it might be affected by other noisy components.

\subsection{On-line normalization algorithm}

We perform an on-line estimation of $H_{01}[e^{j\omega_k}]$ as follows.

Initialization: \\
\begin{align}
  H_u[0, e^{j\omega_k}] = \frac{X_0[0, e^{j\omega_k}]} 
        {X_u[0, e^{j\omega_k}]}  
\end{align}

However, we might need to apply a weighting based on the power.
Let $p[m, e^{\omega_k}]$ represent the power spectrum defined by
$p[m, e^{\omega_k}] = |X[m, e^{\omega_k}]|^2$

If we use the same forgetting factor $\lambda$ for 
$H_{01}[m, e^{j \omega_k}]$ to obtain the smoothed transfer function
$\hat{H}_{01}[m, e^{j \omega_k}]$:
\begin{align}
            \label{eq:h_smoothing}
  \hat{H}_u[m, e^{j\omega_k}] 
      = &  \lambda \hat{H}_u[m - 1, e^{j \omega_k}]  \\ \nonumber 
        &  \qquad + (1 - \lambda) H_{01}[m, e^{j \omega_k}] 
\end{align}

However, the problem with \eqref{eq:h_smoothing} is that
it does not consider the relative intensity at each frame.
More specifically, if the current spectrum at the $m$-th frame is
strong, then s 
of the previous smoothed transfer function $\hat{H}[m, e^{j \omega_k}]$ 
and the current transfer function $H[m, e^{j \omega_k}]$, and 
each of them are weighted by $\lambda$ and $1 - \lambda$, respectively.

The smoothed power $\hat{p}[m, e^{\omega_k}]$ is

We interpret the weighting is, so when the power

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.

\section{Deviation Masking}
After normalizing spectra in the frequency domain, the most straightforward
way is just averaging spectra.
The section in Sec. TODO(chanwcom) based on the assumption that there
is only the target source. If we have both the target and noise source,
the equations becomes as follows:

\begin{subequations}
  \begin{align}
    X_0[m, e^{j\omega_k}]     & = H_0[m, e^{j\omega_k}] X[m, e^{j \omega_k}] + N_0[m, e^{j\omega_k}] \\
                              &  \qquad  \vdots \nonumber \\
    X_u[m, e^{j\omega_k}]     & = H_{u-1}[m, e^{j\omega_k}] X[m, e^{j \omega_k}] + N_u[m, e^{j\omega_k}]\\
                              &  \qquad  \vdots \nonumber \\
      X_{U-1}[m, e^{j\omega_k}] & = H_{U-1}[m, e^{j\omega_k}] X[m, ^{j \omega_k}]+ N_U[m, e^{j\omega_k}]
  \end{align}
  \label{eq:microphone_signals_stft}
\end{subequations}

\section{Experimental Results}
%
TODO(chanwcom) 
% Improve database description and add experimental results.
In this section we describe experimental results obtained
on the DARPA Resource Management 1 (RM1) database and the large scale
English and Icelandic databases used in Google. For the RM1 experiment
we use the Sphinx 3 speech recognition system for acoustic modeling
and decoding, and for the Google speech recognition databases, we
used the state-of-the-art DNN training system, and the GRECO3
decoder. 
TODO(chanwcom)
% Add reference about Google speech recognition system.

\section{Conclusion}
In this paper, we presented an algorithm for microphone compensation.
We have observed that under reverberation, even if the target is placed
at the 
In this paper, we presented an algorithm for phase offset compensation.
Many of the previous algorithms are based on the assumption that if
the target is along the perpendicular bisector to the line connecting
two microphones, the ITD or the phase difference needs to be zero.
However, in the real environments, this assumption usually does not hold.
This discrepancy might occur due to either non-ideal hardware, or
deviation of the target source location.
This POC algorithm can significantly compensate the distortion and
enhance speech recognition accuracy in the real environment.

% use section* for acknowledgement
\section*{Acknowledgements}
This research was supported by Google.

\bibliographystyle{IEEEtran}
\bibliography{../../common_bib_file/common_bib_file}

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}