### 2019-09-24 none_feature_mapping_weve/doc/main.tex
updates the paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
%\usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{tikz,placeins}
\usepackage{subcaption}
%\usepackage[caption = false]{subfig}
\usepackage{amssymb,amsmath,bm}
\usetikzlibrary{arrows,decorations.markings}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\large]

\graphicspath{{../figures/}}

\title{WEVE: Weighted Enhancement Using Variance Estimation}

%\renewcommand{\vec}[1]{\mathbf{#1}}

\newcommand{\chanwcom}{{C. Kim}}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.85}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  David S.~Hippocampus\thanks{Use footnote for providing further
    information about author (webpage, alternative
    address)---\emph{not} for acknowledging funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
TODO(chanw.com) Revise the abstract.
This paper presents a new enhacement algorithm using neural networks 
refered to as WEVE (Weighted Enhancement Using Variance Estimation).
The conventional way of feature enhacement is to directly estimate the clean
feature from the corrupt feature. This mapping from the corrupt feature
to the corresponding clean feature is learned using a training database
consisting of a large number of such pairs of corrupt and clean features
usually with the Mean Squared Error (MSE) loss. The problem of such 
conventional techniques is that they do consider the reliablity of enhancement.
We have frequently observed that this enhancement mapping actually degrades
performance rather than enhacing the classification or recognition
performance under certain conditions.  In WEVE, we estimate both the enhanced 
feature and the variance of this estimation. The final output of the enhancement 
block is a linear combination of the original feature and the 
enhanced feature. The weighting is given by the Minimum Mean
Sqaured Error (MMSE) criterion. This process is very simple nontheless
shows significant improvements in our experiments.
\end{abstract}

\section{Introduction}

\section{Feature enhancement}

In the feature enhancement process, we estimate a clean feature 
$\vec{y}$ given a corrupt feature $\vec{x}$. Suppose that
we have

In this case, the following $L_2$ norm is widely used:
\begin{align}
  l(\vec{y}, \vec{x}) = 
\end{align}

Feature mapping is a process of mapping noisy feature vectors
to clean feature vectors to enhance robustness
\cite{C_Kim_ASRU_2009_1, k_han_taslp_2015_00, k_han_icassp_2014_00}.
This procedure may be formulated by the following equation:
\begin{align}
  Y() = X()
\end{align}

From noisy features, we may infer 
Another

Masking and feature mapping are closely related techniques.
In feature mapping, enhanced features are predicted from
noisy features using machine learning models
The masking-based approaches with
Joint Adaptive Training (JAT) have been shown
to be effective to a large training set
\cite{a_narayanan_interspeech_2015_1}.

NIPS requires electronic submissions.  The electronic submission site
is
\begin{center}
  \url{https://cmt.research.microsoft.com/NIPS2017/}
\end{center}

Please read carefully the instructions below and follow them
faithfully.

\section{The structure of the SPL algorithm}

Fig. \label{fig:entire_diagram} shows the block diagram for
enhancing the noisy feature.

\begin{figure}
    \centering
    \resizebox{80mm}{!}{%
         \input{../figures/entire_diagram.tex}} %
    \caption {
      The structure of feature enhancement using peak
      value normalization and LSTMs.
      \label{fig:entire_diagam}
    }
    %\end{center}
\end{figure}

Fig. \ref{fig:entire_diagram} shows the structure for enhancing
noisy feature.

\subsection{Simulated utterance generation}
We use the room simulation system introduced in
\cite{C_Kim_INTERSPEECH_2017_1} to generate noisy speech
from clean speech.
\begin{align}
  y[n]
\end{align}
For 
\subsection{Review of the Room Simulator for Data Augmentation}
%
%
\begin{figure}
  \centering
    \begin{subfigure}{\linewidth}
      \centering
      \includegraphics[width=70mm]{room_diagram.eps}
      \caption{}
      \label{fig:room_diagram}
    \end{subfigure}
   \\
    \begin{subfigure}{\linewidth}
      \centering
      \includegraphics[width=70mm]{image_method.eps}
      \caption{}
      \label{fig:image_method}
    \end{subfigure}
    \caption{
    (a) A simulated room: There may be multiple microphones, a single target
    sound source, multiple noise sources in a cuboid-shape room with
    acoustically reflective walls. (b) A diagram showing the location
    of the real and the virtual sound sources assuming that the walls
    reflects acoustic wave like a mirror \cite{C_Kim_INTERSPEECH_2017_1}.}
  \vspace{-10mm}
\end{figure}
%
%
In this section, we briefly review the structure of the Google
Room Simulator for generating simulated utterances
to train acoustic models for speech recognition systems
\cite{C_Kim_INTERSPEECH_2017_1}. We assume a room of a
rectangular cuboid-shape as shown in Fig. \ref{fig:room_diagram}.
Assuming that all the walls of a room reflect acoustically uniformly,
we use the image method to model the Room Impulse Responses (RIRs)
\cite{J_Allen_JASA_1979, E_A_Lehmann_ASPAA_2007, S_G_McGovern_RIR}.

Fig. \ref{fig:image_method} shows how virtual sources
appearing as empty circles are constructed in the image method.
Even though the virtual rooms appear to be created in two-dimensions
in Fig. \ref{fig:image_method}, in the ``\emph{Room Simulator}",
they are constructed in three-dimensions. For example, in our
work for training the acoustic model for Google Home, we consider
$17 \times 17 \times 17 = 4913$ virtual rooms for RIR calculation
\cite{C_Kim_INTERSPEECH_2017_1}.
Following the image method, the impulse response is calculated
using the following equation
\cite{ J_Allen_JASA_1979, E_A_Lehmann_ASPAA_2007}:
\begin{align}
    h[n] = \sum_{i = 0}^{I-1} \frac{r^{g_i}}{d_i}
    \delta \left[n -\left \lceil{\frac{d_i f_s}{c_0}}\right \rceil \right],
      \label{eq:h_n_calculation}
\end{align}
where $i$ is the index of each virtual sound source, and $d_i$ is the distance
from that sound source to the microphone, $r$ is
the reflection coefficient of the wall, $g_i$ is
the number of the reflections
to that sound source, $f_s$ is the sampling rate of the RIR, and $c_0$ is the
speed of sound in the air. We use the value of $f_s = 16,000$ \textit{Hz} and
$c_0 = 343$ \textit{m/s} for the room simulator \cite{B_Li_INTERSPEECH_2017_1}.
For $d_i$, $r$, we use numbers created by a random number generator
following specified distributions for each impulse response \cite{C_Kim_INTERSPEECH_2017_1}.

Assuming that there are $I$ sound sources including one target
source and $J$ microphones, the received signal at microphone
$j$ is given by:
\begin{align}
  y_j[n] =  \sum_{i=0}^{I-1} \alpha_{ij} \left(h_{ij}[n] * x_i[n]\right).
  \label{eq:y_j_def}
\end{align}
Since we used a two-microphones system in
\cite{B_Li_INTERSPEECH_2017_1, C_Kim_INTERSPEECH_2017_1},
$J$ is two, and for the number noise sources, we used a value from zero
up to three \cite{C_Kim_INTERSPEECH_2017_1}.
%

\subsection{Feature extraction}

Even though the log-mel coefficients have been widely used
as features for speech recognition \cite{sbdavisandpm1980},
the log non-linearity has a disadvantage in that the value
diverges to negative infinity as the mel filterbank coefficient
approaches zero \cite{C_Kim_IEEETran_2016_1}.
Thus, we use the power-law nonlinearity
rather than the log-nonlinearity:
\begin{align}
  q[m, l] =  (p[m, l])^{\frac{1}{15}}
  \label{eq:power_law_nonlineartiy}
\end{align}
We use the power coefficient of $\frac{1}{15}$ as in
\cite{C_Kim_IEEETran_2016_1, C_Kim_ICASSP_2012_1}.

\subsection{Feature normalization using the peak-value}

The power-mel coefficients $q_x[m, l]$ and $q_y[m, l]$
are obtained from the clean and the simulated noisy
speech $x[n]$ and $y[n]$ respectively. $q_{x, \text{peak}}$
and $q_{y, \text{peak}}$

\section{Estimation of the expected value and the variance
  of the enhanced feature from the noisy feature}
\label{sec:estimation}

\begin{figure}
    \centering
    \resizebox{80mm}{!}{%
         \input{../figures/training.tex}} %
    \caption {
      Training of neural networks to estimate the clean feature and error ratio from corrupt feature
      \label{MFCCPNCCComparison}
    }
    %\end{center}
\end{figure}

In this section, we discuss how to predict the clean feature and the
error ratio from the corrupt feature. Let us denote the clean target feature,
the corrupt feature by $\boldsymbol{x}, \boldsymbol{y}$ respectively.

Using the room simulation system described in \cite{C_Kim_INTERSPEECH_2017_1},
we create a pair of clean and
corrupt utterances. Log-mel [XX] features are calculated from these original clean
and corrupt utterances. Let us denote the clean feature by $\boldsymbol{t}_i$ and
the simulated corrupt feature by $\boldsymbol{x}_i$, respectively where $i$ is the
utterance index. The training set is represented by the following set:
\begin{align}
  \mathcal{T} = \{<\boldsymbol{x}_i, \boldsymbol{t}_i>| 0 \le i \le N - 1\}
\end{align}
where $N$ is the number of training examples.

In the ECE algorithm, we first estimate the true target $\boldsymbol{t}_i$. The
estimation error is defined by the following equation:
\begin{align}
  \boldsymbol{e}_i =  \boldsymbol{t}_i - \boldsymbol{x}_i, \quad 0 \le i \le N - 1
\end{align}

The norm of the error vector $\boldsymbol{e}_i$ would be generally large when
the norm of the estimated clean feature is large.

Thus, instead of trying to estimate the error itself, the neural network tries to
estimate the error ratio given as follows:

\begin{align}
  r_i = {}{}
\end{align}

\section{Generation of simulated clean and noisy data sets}
\label{gen_inst}

To train neural networks described in section \ref{sec:estimation}, we
need a training set consisting of pairs of clean speech and noisy speech.
Unfortunately, it is not easy to have such a training set from real speech
utterances. Thus, we use the simulation system described in
\cite{C_Kim_INTERSPEECH_2017_1} to synthetically generate noisy speech
utterances from the clean speech utterances.

In our application, for clean training set, we used Wall Street Journal (WSJ)
si-284. For noise set, we used the Resource Management 1 (RM1)
\cite{price_p_ldc_1993} 50 \% of time 
and DEMAND noise [TODO(chanwcom) Adds reference] for the remaining 50 \% of time.

TODO(chanwcom) Cite DEMAND noise set.

\section{Experiments}

MNIST experiment
RMDL experiment \cite{k_kowsari_icisdm_2018_00} 
\cite{c_wei_bmvc_2018_00}

%
%
%
%\begin{figure*}
%  \centering
%    \begin{subfigure}[\label{fig:randomly_selected_grid_coordinates}]{0.5\textwidth} {
%        \includegraphics[width=\textwidth]{target_estimation}}
%    \caption{Grid from random-sampling}
%    \end{subfigure}
%   % \begin{subfigure}[\label{fig:intersection_point}]{0.5\textwidth} {
%   %   \includegraphics[width=\textwidth]{intersection_point}}
%   % \caption{Calculation of grid coordinates}
%   % \end{subfigure}
%  \caption {Random Grid Sampling}
%\end{figure*}
%
%
%
\section{Estimation of the optimal weight in the EFW algorithm}
In this section, we describe the procedure of feature weighting approach.
Suppose that the feature before enhancement is represented by $\vec{x}$.
Let us represent the inference neural network to estimate the target by
$\mathcal{T}$.

\begin{align}
  \mathcal{X} = \{  \}
\end{align}

Instead of directly using the enhanced feature $\widehat{\vec{y}}$, we consider
the following interpolated feature $\bm{z}$ from the original corrupt
input $\bm{x}$ and .
\begin{align}
  \vec{z} = \vec{w} \odot \widehat{\vec{y}} \ + (1 - \vec{w}) \odot \vec{x} 
    \label{eq:vec_z}
\end{align}
where $\odot$ denotes the Hadamard product (entry-wise product).

Let us denote the estimated variance vector by $\hat{\vec{v}}$:
\begin{align}
  \hat{\vec{v}} & = \text{Var}[\widehat{\vec{y}}] \nonumber \\
                & = E[(y - y)^2 | \vec{x}[0], \vec{x}[1], ... \vec{x}[T]]
     \label{eq:vec_v}
\end{align}

In our discussion, let us assume that the expectation of $\hat{\vec{y}}$ is the
same as the true target $\vec{y}$.
\begin{align}
  E[\widehat{\vec{y}}] = \vec{y}  \label{eq:expectation_of_y}
\end{align}

Now, let us obtain the expected value of $\vec{z}$ from
\eqref{eq:vec_z} and \eqref{eq:expectation_of_y}
\begin{align}
  E[\vec{z}] = (1 - \vec{w}) \odot (\vec{x} - \vec{y})
\end{align}
Thus, the bias of $\vec{z}$ is given by:
\begin{align}
  Bias_{\vec{z}} & = E[\vec{z}] - t \nonumber \\
                 & = \vec{w} \odot (\widehat{\vec{y}} - \vec{y})
\end{align}

From \eqref{eq:vec_z} and \eqref{eq:vec_v}, the variance of $\vec{z}$ is given by:
\begin{align}
  Var_{\vec{z}} = \vec{v} \odot w \odot w.
\end{align}
The mean squared error of the $i$-th element of the $\vec{z}$ is given by:
\begin{align}
  MSE_{\vec{z_i}} & = Bias_{\vec{z_i}}^2 + Var_{\vec{z_i}} \nonumber \\
                 & =  \left[(\vec{x}_i - \vec{y}_i) ^ 2
                    + \vec{v}_i^2 \right] \vec{w}_i^{2}
                    - 2 (\vec{x}_i - \vec{y}_i)^{2}  \vec{w}_i
                    + (\vec{x}_i - \vec{y}_i)^{2}
 \label{eq:mse_z_i}
\end{align}
Since the above equation \eqref{eq:mse_z_i} is a quadratic equation with
respect to $\vec{w}_i$, we may find that the minimum value of $MSE_{\vec{z_i}}$
is obtained when $\vec{w}_i$ has the following value:
\begin{align}
  \vec{w}_i = \frac{(\vec{x}_i - \vec{y}_i)^2}{(\vec{x}_i - \vec{y}_i)^2 + \vec{v}_i}
\end{align}

Thus, the final form of the estimated vector $\widehat{\vec{z}}$ becomes:
\begin{align}
  \widehat{\vec{z}} = \widehat{\vec{w}} \odot \widehat{\vec{y}}
    \ + (1 - \widehat{\vec{w}}) \odot \vec{x}
\end{align}
where
\begin{align}
  \widehat{\vec{w}} = {(\vec{x} - \vec{y})^{\odot 2}}
    \left[(\vec{x} - \vec{y})^{\odot 2}+ \vec{v}\right]^{\odot -1}.
\end{align}

Sequence enhancement

The feature enhancement is to map a sequence of corrupt features $\mathbf{X}$.
\begin{subequations}
  \begin{align}
    \mathbf{X} = \left(\vec{x}[0], \vec{x}[1], \cdots, \vec{x}[M-1] \right)  \\
    \mathbf{Y} = \left(\vec{y}[0], \vec{y}[1], \cdots, \vec{y}[M-1] \right) 
  \end{align}
\end{subequations}

\subsubsection*{Acknowledgments}

\section*{References}

References follow the acknowledgments. Use unnumbered first-level
heading for the references. Any choice of citation style is acceptable
as long as you are consistent. It is permissible to reduce the font
size to \verb+small+ (9 point) when listing the references. {\bf
  Remember that you can go over 8 pages as long as the subsequent ones contain
  \emph{only} cited references.}
\medskip

\small

\clearpage
\newpage
\bibliographystyle{plain}
\bibliography{../../common_bib_file/common_bib_file}

\end{document}