### 2019-12-15 asru_2019_maximizing_uniformity_of_distribution/poster/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% a0poster Landscape Poster
% LaTeX Template
% Version 1.0 (22/06/13)
%
% The a0poster class was created by:
% Gerlinde Kettl and Matthias Weiser (tex@kettl.de)
% 
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0,portrait]{a0poster}
%\documentclass[a4]{article}

\usepackage{multicol} % This is so we can have multiple columns of text side-by-side
\columnsep=100pt % This is the amount of white space between the columns in the poster
\columnseprule=3pt % This is the thickness of the black line between the columns in the poster

\usepackage[svgnames]{xcolor} % Specify colors by their 'svgnames', for a full list of all colors available see here: http://www.latextemplates.com/svgnames-colors

\usepackage{times} % Use the times font
%\usepackage{palatino} % Uncomment to use the Palatino font

\usepackage{float}
\usepackage{graphicx} % Required for including images
\usepackage{booktabs} % Top and bottom rules for table
\usepackage[export]{adjustbox}
\usepackage[font=large,labelfont=bf]{caption} % Required for specifying captions to tables and figures
%\usepackage{tikzpagenodes}
\usepackage{amsfonts, amsmath, amsthm, amssymb, bm} % For math fonts, symbols and environments
\usepackage{wrapfig} % Allows wrapping text around tables and figures
\usepackage{multirow}
\usepackage[justification=centering]{caption}
\usepackage[caption = false]{subfig}

\DeclareMathOperator*{\argmax}{arg\,max}

\usepackage{tikz,placeins}
\usetikzlibrary{arrows,decorations.markings}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\tiny]
\usetikzlibrary{positioning, shapes}
%\usepackage{xcolor} %already loaded by tikz
%
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\graphicspath{{../figures/}}

\newcommand{\specialcell}[2][c]{%
     \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newcommand{\chanwcom}{{C. Kim}}

\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

\begin{document}

%----------------------------------------------------------------------------------------
%	POSTER HEADER 
%----------------------------------------------------------------------------------------

% The header is divided into three boxes:
% The first is 55% wide and houses the title, subtitle, names and university/organization
% The second is 25% wide and houses contact information
% The third is 19% wide and houses a logo for your university/organization or a photo of you
% The widths of these boxes can be easily edited to accommodate your content as you see fit

\begin{minipage}[t]{1.0\linewidth}
\begin{center}
  \begin{minipage}[t]{0.1\textwidth}
    \centering
    \vspace{0pt}
    \includegraphics[width=10cm, valign=t]{samsung_research} % Logo or a photo of you, adjust its dimensions here
  \end{minipage} \hfill
  \begin{minipage}[t]{0.7\textwidth}
    \begin{center}
  \huge \color{NavyBlue} 
      \textbf{Power-law nonlinearity with maximally uniform distribution criterion for
      improved neural network training in automatic speech recognition} \\
      \LARGE \textbf{Chanwoo Kim, Mehul Kumar, Kwangyoun Kim, and Dhananjaya Gowda}
  \\
\LARGE Samsung Research, Seoul, South Korea\\ % University/organization
\large \{chanw.com, mk0211.shin, ky85.kim, d.gowda\}@samsung.com
    \end{center}
  \end{minipage}
  \begin{minipage}[t]{0.1\textwidth}
    \begin{center}
    \includegraphics[width=10cm]{samsung_research} % Logo or a photo of you, adjust its dimensions here
    \end{center}
  \end{minipage} \hfill
\end{center}
\end{minipage}
%
%\begin{minipage}[b]{0.25\linewidth}
%\color{DarkSlateGray}\Large \textbf{Contact Information:}\\
%Department Name\\ % Address
%University Name\\
%123 Broadway, State, Country\\\\
%Phone: +1 (000) 111 1111\\ % Phone number
%Email: \texttt{john@LaTeXTemplates.com}\\ % Email address
%\end{minipage}
%

\vspace{1cm} % A bit of extra whitespace between the header and poster content

%----------------------------------------------------------------------------------------

\begin{multicols}{2} % This is how many columns your poster will be broken into, a poster with many figures may benefit from less columns whereas a text-heavy poster benefits from more

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\large

\color{Navy} % Navy color for the abstract

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\color{SaddleBrown} % SaddleBrown color for the introduction

\section{Motivation}

\begin{enumerate}
  \item Applies Vocal Tract Length Perturbation (VTLP) to diversify the
    training database \cite{n_jaitly_icml_workshop_2013_00}.
  \item Resynthesizes VTLP-processed waveforms before performing feature extraction.
    \begin{itemize}
        \item We can apply the acoustic simulator after performing the VTLP
          processing.
        \item We may use a different window length for the frequency warping from
          the one used in feature processing for performance optimization.
    \end{itemize}
\end{enumerate}

\color{DarkSlateGray} % DarkSlateGray color for the rest of the content
  \section{Maximization of Distribution Uniformity}

\begin{Figure}
  \begin{center}
    {
      \includegraphics[width=170mm]{mel_filterbank_energy_histogram}
      \label{fig:mel_filterbank_energy_histogram}
    }   
    \\  
    {
      \includegraphics[width=170mm]{power_function_based_mud_output_histogram}
      \label{fig:power_function_based_mud_output_histogram}
    }   
  \vspace{-2mm}
    \\  
    {
      \includegraphics[width=170mm]{histogram_based_mud_output_histogram}
      \label{fig:histogram_based_mud_output_histogram}
    }   
  \vspace{-2mm}
    \captionof{figure} {
      The Probability Density Functions for the third filterbank channel
       of (a): the mel filterbank energy $p[m, l], l = 3$ in  
       \eqref{eq:mel_filterbank_energy}, (b): the power-function based MUD 
      output of this mel filterbank energy in \eqref{eq:power_mud}, (c) 
      the histogram based-MUD in \eqref{eq:histogram_mud}. 
      \label{fig:pdf_mud}
    }   
   \vspace{-7mm}
  \end{center}
\end{Figure}

%\begin{center}
%  \includegraphics[width=0.2\linewidth]{power_function_based_mud_output_histogram}

%\captionof{figure}{
%  \label{fig:bilinear_transform}
%      The relation between the input frequency and the
%      warped frequency using the bilinear transformation
%      \eqref{eq:bilinear_transformation}.}
%
%\end{center}

%\item \textbf{IFFT and OverLap Addition (OLA)} Re-synthesize the waveform.
%\end{enumerate}

%\begin{figure}%[!tbp]
%\end{figure}

\section{End-to-End Speech recognition}

\begin{minipage}{1.0\linewidth}
  {
    \normalsize
  \begin{center}
  \resizebox{180mm}{!}{\input{../figures/e2e_asr_block_diagram_poster.tex}}
  \end{center}
\captionof{figure}{ 
  The structure of the entire end-to-end speech recognition system.}
  }
\end{minipage}

  We used the RETURNN speech recognition system
  \cite{a_zeyer_interspeech_2018_00} with various modifications:
    \begin{itemize}
      \item{Gradient clipping and learning-late warm-up.}
      \item{Power-mel feature} Motivated by the power-law nonlinearity of ($(\cdot)^{\frac{1}{15}}$) in PNCC \cite{C_Kim_IEEETran_2016_1}.
      \item{Modified shallow fusion with a Transformer LM}

        \begin{align}
  y_{0:L}^{*} = & \argmax_{y_{0:L}} \sum_{l=0}^{L-1} \Big[ \log P \left(y_l
  |\vec{x}[0:M], y_{0:l} \right)
  \nonumber \\
                           & \qquad - \lambda_p \log P(y_l)
                           + \lambda_{\text{lm}} \log P \left(y_l   |y_{0:l} \right)  \Big],
                           \label{eq:lm_fusion}
\end{align}

    \end{itemize}

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	MATERIALS AND METHODS
%----------------------------------------------------------------------------------------

\section{Experimental Results}
\begin{itemize}
  \item For training, we used { \bf the Librispeech train-clean and
    train-other}.
  \item For evaluation, we used { \bf the test-clean and test-other of this Libri
    Speech database. }
\end{itemize}

\begin{minipage}[b]{1.0\linewidth}
  %\begin{table*}[!htbp]
  \renewcommand{\arraystretch}{1.3}
  \centering
      \captionof{table}{\label{tbl:result}
      Word Error Rates (WERs) obtained with MFCC, Power Mel filterbank
      coefficients,  power function-based MUD processing, and
      histogram-based MUD Processing on the LibriSpeech corpus
      \cite{v_panayotov_icassp_2015_00}. 
      }
      %\vspace{-3mm}
      \begin{tabular}{| c | c  || c | c | c | c |}
        \hline
        \multicolumn{2}{| l ||}{Neural Network Structure}
                               & \specialcell{MFCC}
                               & \specialcell{$ ( \cdot ) ^ {\frac{1}{15}}$}
                               & \specialcell{Power Function- \\ Based MUD}
                               & \specialcell{Histogram- \\ Based MUD}   \\
        \hline \hline
        \multirow{3}{*}{\specialcell{1024 cell \\ ULSTM \\ MoCha}}
                & test-clean  &   7.09  \% & \textcolor{blue}{\textbf{ 7.04 }} \%   &   7.10 \%  &  7.13 \%  \\
                & test-other  &  20.60  \% & 19.76 \%
                &  \textcolor{blue}{\textbf{19.64}} \%  & 20.03 \%  \\
                & average     &  13.85  \% & 13.40 \%
                &  \textcolor{blue}{\textbf{13.37}} \%  & 13.58 \%  \\
        \hline \hline

            \multirow{3}{*}{\specialcell{1536 cell \\ BLSTM \\ Full-Attention}}
                & test-clean  &   4.06  \% & \textcolor{blue}{\textbf{3.94}} \%   &   4.02 \%  &  4.11 \%  \\
                & test-other  &  13.97  \% & 13.56 \%
                & \textcolor{blue}{\textbf{13.34}} \%  & 14.10 \%  \\
                & average     &   9.02  \% &  8.75 \%
                & \textcolor{blue}{\textbf{8.68}} \%  &  9.11 \%  \\
        \hline
     \end{tabular}
     \vspace{-4mm}
\end{minipage}
 %----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

%\nocite{*} % Print all references regardless of whether they were cited in the poster or not
%\bibliographystyle{plain} % Plain referencing style
%\bibliography{sample} % Use the example bibliography file sample.bib

%
\color{SaddleBrown} % SaddleBrown color for the introduction
\section{Conclusions}
%
{ \color{red}
We achieved 2.44 \% WER and 8.29 \% WER
  on the Librispeech {\it test-clean} and {\it test-other} databases.
To the best of our knowledge, the 2.44 \% WER on the {\it test-clean}
is the best result ever reported on this database
}
%
%
\begin{enumerate}
  \item Applies Vocal Tract Length Perturbation (VTLP) to diversify the
    training database. \cite{n_jaitly_icml_workshop_2013_00}
  \item Resynthesizes VTLP-processed waveforms before performing feature extraction.
    \begin{itemize}
        \item We can apply the acoustic simulator after performing the VTLP
          processing.
        \item We may use a different window length (50 {\it ms}) for the frequency warping from
          the one used in feature processing (25 {\it ms}).
    \end{itemize}
  \item Applies the power-mel feature.
  \item Applies a modified shallow-fusion with a Transformer LM.
\end{enumerate}
\color{DarkSlateGray} % DarkSlateGray color for the rest of the content
%----------------------------------------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{../../common_bib_file/common_bib_file}
\end{multicols}
\end{document}