### 2019-12-16 asru_2019_training_infra/poster/main.tex
Update the poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% a0poster Landscape Poster
% LaTeX Template
% Version 1.0 (22/06/13)
%
% The a0poster class was created by:
% Gerlinde Kettl and Matthias Weiser (tex@kettl.de)
% 
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0,portrait]{a0poster}
%\documentclass[a4]{article}

\usepackage{multicol} % This is so we can have multiple columns of text side-by-side
\columnsep=100pt % This is the amount of white space between the columns in the poster
\columnseprule=3pt % This is the thickness of the black line between the columns in the poster

\usepackage[svgnames]{xcolor} % Specify colors by their 'svgnames', for a full list of all colors available see here: http://www.latextemplates.com/svgnames-colors

\usepackage{times} % Use the times font
%\usepackage{palatino} % Uncomment to use the Palatino font

\usepackage{float}
\usepackage{graphicx} % Required for including images
\usepackage{booktabs} % Top and bottom rules for table
\usepackage[export]{adjustbox}
\usepackage[font=large,labelfont=bf]{caption} % Required for specifying captions to tables and figures
%\usepackage{tikzpagenodes}
\usepackage{amsfonts, amsmath, amsthm, amssymb, bm} % For math fonts, symbols and environments
\usepackage{wrapfig} % Allows wrapping text around tables and figures
\usepackage{multirow}
\usepackage[justification=centering]{caption}
\usepackage[caption = false]{subfig}

\DeclareMathOperator*{\argmax}{arg\,max}

\usepackage{tikz,placeins}
\usetikzlibrary{arrows,decorations.markings}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\tiny]
\usetikzlibrary{positioning, shapes}
%\usepackage{xcolor} %already loaded by tikz
%
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\graphicspath{{../figures/}}

\newcommand{\specialcell}[2][c]{%
     \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newcommand{\chanwcom}{{C. Kim}}

\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

\begin{document}

%----------------------------------------------------------------------------------------
%	POSTER HEADER 
%----------------------------------------------------------------------------------------

% The header is divided into three boxes:
% The first is 55% wide and houses the title, subtitle, names and university/organization
% The second is 25% wide and houses contact information
% The third is 19% wide and houses a logo for your university/organization or a photo of you
% The widths of these boxes can be easily edited to accommodate your content as you see fit

\begin{minipage}[t]{1.0\linewidth}
\begin{center}
  \begin{minipage}[t]{0.1\textwidth}
    \centering
    \vspace{0pt}
    \includegraphics[width=10cm, valign=t]{samsung_research} % Logo or a photo of you, adjust its dimensions here
  \end{minipage} \hfill
  \begin{minipage}[t]{0.79\textwidth}
    \begin{center}
  \huge \color{NavyBlue} 
      \textbf{End-to-end Training of a Large Vocabulary \\ End-to-end Speech Recognition System
      } \\
      \Large \textbf{
      Chanwoo Kim, Sungsoo Kim, Kwangyoun Kim, Mehul Kumar,  Jiyeon Kim, \\ 
      Kyungmin Lee, Changwoo Han, Abhinav Garg, Eunhyang Kim,  \\ 
      Minkyoo Shin, Shatrughan Singh, Larry Heck, Dhananjaya Gowda
      }
  \\
\Large Samsung Research, Seoul, South Korea\\ % University/organization
{ \large \tt  \{chanw.com, ss216.kim, ky85.kim, mehul3.kumar, jstacey7.kim,} \\ 
  \large \tt  {k.m.lee, cw1105.han, abhinav.garg, sc.ehkim.jin,} \\
  \large \tt  {mk0211.shin, shatrughan.s, larry.h, d.gowda\}@samsung.com }
    \end{center}
  \end{minipage}
  \begin{minipage}[t]{0.1\textwidth}
    \begin{center}
    \includegraphics[width=10cm]{samsung_research} % Logo or a photo of you, adjust its dimensions here
    \end{center}
  \end{minipage} \hfill
\end{center}
\end{minipage}
%
%\begin{minipage}[b]{0.25\linewidth}
%\color{DarkSlateGray}\Large \textbf{Contact Information:}\\
%Department Name\\ % Address
%University Name\\
%123 Broadway, State, Country\\\\
%Phone: +1 (000) 111 1111\\ % Phone number
%Email: \texttt{john@LaTeXTemplates.com}\\ % Email address
%\end{minipage}
%
%
\vspace{-2mm} % A bit of extra whitespace between the header and poster content
%
%----------------------------------------------------------------------------------------
%
\begin{multicols}{2} % This is how many columns your poster will be broken into, a poster with many figures may benefit from less columns whereas a text-heavy poster benefits from more
%
%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
%
%
\large
%
\color{Navy} % Navy color for the abstract
%
%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------
%
\color{SaddleBrown} % SaddleBrown color for the introduction
%
\section{Summary}
%
\begin{itemize}
\item Achieves the state of the art speech recognition performance for commercial products using an end-to-end pipeline that performs all of data reading,  large scale data augmentation \cite{c_kim_interspeech_2019_00}, power-mel feature extraction \cite{c_kim_asru_2019_00}, and  distributed neural network parameter updates in an “on-the-fly” way.
\item Performs Vocal Tract Length Perturbation and Acoustic Simulation \cite{C_Kim_INTERSPEECH_2017_1} on the CPU queue using example servers.
\item Performs Neural Beamforming \cite{j_heymann_icassp_2016_00} on the device side for further improvement in far-field noisy environments.
\item Performs experiments both using the on-line MoCha \cite{k_kim_asru_2019_00} and the Bidirectional Full-Attention (BFA) approach.
\end{itemize}
%
%
\color{DarkSlateGray} % DarkSlateGray color for the rest of the content
  \section{Overall Structure}
The overall structure is shown in Fig. \ref{fig:end_to_end_training_system}
\begin{itemize}
\item \textbf{Data reading}: Using sharded {\tt TFRecords} and {\tt tf.data}.
\item \textbf{Data augmentation and feature extraction}: VTLP, AS, and power-mel feature extraction are running on the example servers.
\item \textbf{Connection between the example servers and the GPU servers}: Using {\tt ZeroMQ} for asynchronous message queueing.
\item \textbf{Distributed neural net training}: Training on the GPU server using {\tt horovod}.
\end{itemize}
%
%
\begin{Figure}
  \begin{center}
    \resizebox{\textwidth}{!}{\input{../figures/end_to_end_training_system_poster.tex}}
      \captionof{figure} { The Samsung Research end-to-end training framework for building an end-to-end
  speech recognition system with multi CPU-GPU clusters and on-the-fly data
  processing and augmentation pipeline.
     }
     \label{fig:end_to_end_training_system}
  \end{center}
\vspace{-5mm}
\end{Figure}
%
%
%
%
\begin{minipage}[]{0.45\linewidth}
  {
    \normalsize
  \begin{center}
  \resizebox{120mm}{!}{\input{../figures/e2e_asr_block_diagram_poster.tex}}
  \end{center}
\captionof{figure}{ 
  The neural network structure for end-to-end speech recognition.}
  }
\end{minipage}
%
%
\begin{minipage}[]{0.54\linewidth}
  We used the RETURNN speech recognition system \cite{a_zeyer_interspeech_2018_00} with various modifications:
    \begin{itemize}
      \item{MoCha\cite{k_kim_asru_2019_00} and the modified beam search decoder.}
      \item{Gradient clipping, modified learning-late warm-up, and so on.}
      \item{Power-mel feature} Motivated by the power-law nonlinearity of ($(\cdot)^{\frac{1}{15}}$).
      \item{Modified shallow fusion with a Transformer LM \cite{c_kim_interspeech_2019_00}}
        \begin{align}
  y_{0:L}^{*} = & \argmax_{y_{0:L}} \sum_{l=0}^{L-1} \Big[ \log P \left(y_l
  |\vec{x}[0:M], y_{0:l} \right)
  \nonumber \\
                           & \qquad - \lambda_p \log P(y_l)
                           + \lambda_{\text{lm}} \log P \left(y_l   |y_{0:l} \right)  \Big],
                           \label{eq:lm_fusion}
\end{align}
    \item{Example queues for efficient data augmentation}
    \end{itemize}
\end{minipage}
%
%
\begin{Figure}
  \captionsetup[subfigure]{justification=centering}
    \centering
    \begin{minipage}{0.49\linewidth}
			\centering
      {\includegraphics[width=175mm]{../figures/plot_example_server_time}  \label{fig:plot_example_server_time} }
			{(a)}	
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
			\centering
      {\includegraphics[width=175mm]{../figures/plot_gpu_time_percentage} \label{fig:plot_gpu_time_percentage} }
			{(b)}	
    \end{minipage}
  \captionof{figure} {\label{fig:plot_example_server_performance}
  The efficiency of the {\it example server} with respect to
  the number of CPUs per GPU: (a) The required time to
  process a single epoch during the training phase and (b)
  the percentage of {\tt Tensorflow } computation time defined by
  $t_{\text{session}} =  \frac{\text{Time Spent in {\tt Tensorflow}
  Session}}{\text{Elapsed Time}}$.
  The blue horizontal dotted lines in each figure represent the case
  when data augmentation with example servers is not employed.
  }
\end{Figure}
%
%
\section{Experimental Results}
%
%
%
%
\begin{itemize}
\item \textbf{LibriSpeech experiments} (960-hr training and evaluation on 5.4-hr LibriSpeech {\tt test-clean}) \\
\end{itemize}
\begin{minipage}[b]{1.0\linewidth}
%\begin{table}[!tbhp]
  \renewcommand{\arraystretch}{1.3}
  \centering
        \captionof{table}{\label{tbl:near_field_result}
        Summary of Word Error Rates (WERs) obtained for different
        LibriSpeech and {\tt Bixby } near-field end-to-end ASR models
        with and without an RNN LM.
        }
        %\vspace{-3mm}
        \begin{tabular}{| c | c  || c | c |}
          \hline
          \multicolumn{2}{| l ||}{Models}
                                 & BFA
                                 & MoChA \\
          \hline \hline
          \multirow{3}{*}{\specialcell{LibriSpeech \\ (1536-cell) \\ test-clean}}
                  & w/o LM  &   3.66  \% &  6.78  \%  \\
                  & RNN-LM  &   2.85  \% &  5.54  \%  \\
                  & Transformer LM  &   2.44  \% &   -   \\
          \hline
          \multirow{2}{*}{\specialcell{Bixby \\ (1024-cell) }}
                  & w/o LM  &   8.25  \% &  10.77  \% \\
                  & RNN-LM  &   7.92  \% &   9.95  \% \\
          \hline
       \end{tabular}
       \vspace{-2mm}
%\end{table}
\end{minipage}
%
%
%
%
\begin{itemize}
\item \textbf{Bixby English command-set experiments} (10,000-hr {\tt Bixby } training set and {\tt Bixby} command test sets) \\
\item Note that the MoCha version of the system trained using  {\tt Bixby } training set was commercialized for on-device dictation for high-end Samsung mobile phones. \\
\end{itemize}
\begin{Figure}
  \captionsetup[subfigure]{justification=centering}
    \centering
    \begin{minipage}{0.49\textwidth}
			\centering
      {\includegraphics[width=175mm]{../figures/plot_farfield_wer_tv}}
			{(a)}	
    \end{minipage}
%
    \begin{minipage}{0.49\textwidth}
			\centering
      {\includegraphics[width=175mm]{../figures/plot_farfield_wer_music}}
		{(b)}
    \end{minipage}
%
%
  %\vspace{-6mm}
 \captionof{figure} {\label{fig:plot_farfield_wer}
  Speech recognition accuracy at different Signal-to-Noise Ratios (SNRs)
  under three different noisy conditions: direct TV noise (a)
  , music noise (b), and babble noise (c). NBF, VTLP, and AS
  stand for Neural Beam Former (NBF) \cite{j_heymann_icassp_2016_00},
  Vocal Tract Length Perturbation (VTLP) \cite{c_kim_interspeech_2019_00} , and
  Acoustics Simulator (AS) \cite{c_kim_interspeech_2017_00},
  respectively.
  \vspace{-2mm}
 }
\end{Figure}
 %----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

%\nocite{*} % Print all references regardless of whether they were cited in the poster or not
%\bibliographystyle{plain} % Plain referencing style
%\bibliography{sample} % Use the example bibliography file sample.bib

%
\color{SaddleBrown} % SaddleBrown color for the introduction
%\section{Conclusions}
%%
%\begin{enumerate}
%\item We achieved 2.44 \% WER and 8.29 \% WER
%  on the Librispeech {\it test-clean} and {\it test-other} databases.
%%
%\end{enumerate}
%
\color{DarkSlateGray} % DarkSlateGray color for the rest of the content
%----------------------------------------------------------------------------------------
\small
\bibliographystyle{IEEEtran}
\bibliography{../../common_bib_file/common_bib_file}
\end{multicols}
\end{document}