### 2019-12-17 asru_2019_maximizing_uniformity_of_distribution/poster/main.tex
Update the paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% a0poster Landscape Poster
% LaTeX Template
% Version 1.0 (22/06/13)
%
% The a0poster class was created by:
% Gerlinde Kettl and Matthias Weiser (tex@kettl.de)
% 
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0,portrait]{a0poster}
%\documentclass[a4]{article}

\usepackage{multicol} % This is so we can have multiple columns of text side-by-side
\columnsep=100pt % This is the amount of white space between the columns in the poster
\columnseprule=3pt % This is the thickness of the black line between the columns in the poster

\usepackage[svgnames]{xcolor} % Specify colors by their 'svgnames', for a full list of all colors available see here: http://www.latextemplates.com/svgnames-colors

\usepackage{times} % Use the times font
%\usepackage{palatino} % Uncomment to use the Palatino font

\usepackage{float}
\usepackage{graphicx} % Required for including images
\usepackage{booktabs} % Top and bottom rules for table
\usepackage[export]{adjustbox}
\usepackage[font=large,labelfont=bf]{caption} % Required for specifying captions to tables and figures
%\usepackage{tikzpagenodes}
\usepackage{amsfonts, amsmath, amsthm, amssymb, bm} % For math fonts, symbols and environments
\usepackage{wrapfig} % Allows wrapping text around tables and figures
\usepackage{multirow}
\usepackage[justification=centering]{caption}
\usepackage[caption = false]{subfig}

\DeclareMathOperator*{\argmax}{arg\,max}

\usepackage{tikz,placeins}
\usetikzlibrary{arrows,decorations.markings}
\tikzstyle{every picture}+=[font=\rmfamily\it\bfseries\tiny]
\usetikzlibrary{positioning, shapes}
%\usepackage{xcolor} %already loaded by tikz
%
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\graphicspath{{../figures/}}

\newcommand{\specialcell}[2][c]{%
     \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newcommand{\chanwcom}{{C. Kim}}

\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

\begin{document}

%----------------------------------------------------------------------------------------
%	POSTER HEADER 
%----------------------------------------------------------------------------------------

% The header is divided into three boxes:
% The first is 55% wide and houses the title, subtitle, names and university/organization
% The second is 25% wide and houses contact information
% The third is 19% wide and houses a logo for your university/organization or a photo of you
% The widths of these boxes can be easily edited to accommodate your content as you see fit

\begin{minipage}[t]{1.0\linewidth}
\begin{center}
  \begin{minipage}[t]{0.1\textwidth}
    \centering
    \vspace{0pt}
    \includegraphics[width=10cm, valign=t]{samsung_research} % Logo or a photo of you, adjust its dimensions here
  \end{minipage} \hfill
  \begin{minipage}[t]{0.7\textwidth}
    \begin{center}
  \huge \color{NavyBlue} 
      \textbf{Power-law nonlinearity with maximally uniform distribution criterion for
      improved neural network training in automatic speech recognition} \\
      \LARGE \textbf{Chanwoo Kim, Mehul Kumar, Kwangyoun Kim, and Dhananjaya Gowda}
  \\
\LARGE Samsung Research, Seoul, South Korea\\ % University/organization
\large \{chanw.com, mehul3.kumar, ky85.kim, d.gowda\}@samsung.com
    \end{center}
  \end{minipage}
  \begin{minipage}[t]{0.1\textwidth}
    \begin{center}
    \includegraphics[width=10cm]{samsung_research} % Logo or a photo of you, adjust its dimensions here
    \end{center}
  \end{minipage} \hfill
\end{center}
\end{minipage}
%
%\begin{minipage}[b]{0.25\linewidth}
%\color{DarkSlateGray}\Large \textbf{Contact Information:}\\
%Department Name\\ % Address
%University Name\\
%123 Broadway, State, Country\\\\
%Phone: +1 (000) 111 1111\\ % Phone number
%Email: \texttt{john@LaTeXTemplates.com}\\ % Email address
%\end{minipage}
%

\vspace{1cm} % A bit of extra whitespace between the header and poster content

%----------------------------------------------------------------------------------------

\begin{multicols}{2} % This is how many columns your poster will be broken into, a poster with many figures may benefit from less columns whereas a text-heavy poster benefits from more

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\large

\color{Navy} % Navy color for the abstract

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\color{SaddleBrown} % SaddleBrown color for the introduction

\section{Summary about Maximum Uniformity of Distribution (MUD)}
%
\begin{itemize}
  \item We developed a novel approach for obtaining {the \color{Navy} nonlinearity function} in a {\it data-driven } way.
  \item This approach is based on the assumption that neural-network training would be easier 
  and the converged parameters would show better
  performance when feature distribution is not too much
  skewed or too much concentrated in an extremely narrow interval.

  \item The criterion is to maximize {\color{Navy} the Uniformity of Distribution}.
  \item We develop two different types of MUD. 
    \begin{itemize}
      \item Power Function-Based MUD
      \item Histogram-Based MUD
    \end{itemize}
  \item Power function-based MUD shows better results than the histogram-based MUD.
  \item We achieved 4.02 \%  WER on test-clean and 13.34 \% WER on test-other without using any Language Models (LMs).
  \item This approach may be extended to other domains as well.
\end{itemize}
%
%
\color{DarkSlateGray} % DarkSlateGray color for the rest of the content
  \section{Maximization of Distribution Uniformity}

\begin{itemize}
  \item The MUD nonlinearity is applied to mel-filterbank energy for each channel.
\end{itemize}

\subsection{Power-function based maximization of distribution
uniformity}
%
\begin{itemize}
\item The transformation based on the power function:
\begin{align}
  \mathbf{Y} = \sigma_{p} \left(\mathbf{X} \right) = (X - x_{\text{min}})^{\alpha}.
    \label{eq:power_mud}
\end{align}
%
\item We expect $Y$ to follow the following uniform distribution:
\begin{align}
 \mathbf{Y} \sim 
  \mathcal{U}(0, \left(x_{\text{max}} - x_{\text{min}} \right)^{\alpha} ).
    \label{eq:pdf_y}
\end{align}
%
\item
The log likelihood of the data $X$ assuming the PDF in \eqref{eq:pdf_y}
is given by (Refer to the paper for mathematical details):
%
%
\begin{align}
  \mathcal{L}\left(\alpha ; X\right) 
      & = \sum_{i=0}^{N-1} \ln p_{\mathbf{X}} \left( x_i \right)  \nonumber \\
      & = \sum_{i=0}^{N-1} 
              \ln \left[ \frac{ \alpha (x_i - x_{\text{min}} )^{\alpha-1}}  
              {\left(x_{\text{max}} - x_{\text{min}} \right)^{\alpha}} \right]  \nonumber \\
      & = N \ln(\alpha) + \left(\alpha-1 \right) \sum_{i=0}^{N-1}
        \ln \left(x_i - x_{\text{min}} \right)  
       - N \alpha \ln \left(x_{\text{max}} - x_{\text{min}} \right).
        \label{eq:x_likelihood}
\end{align}
%
%
%
%
%
\item 
%
The following $\hat{\alpha}$ maximizes the log-likelihood of $\mathcal{L}\left(\alpha ; X\right)$
(Refer to the paper for mathematical details).
%
%
\begin{align}
  \hat{\alpha} = \frac{1}{
    \ln \left(\max \left\{x_i - x_{\text{min}}, \delta \right\} \right)
    - \frac{1}{N} \sum_{i=0}^{N-1} \ln \left(x_i - x_{\text{min}}\right)}.
  \label{eq:alpha_estimation}
\end{align}
%
%
\end{itemize}
%
%
  \begin{Figure}
    \begin{center}
      \includegraphics[width=170mm]{estimated_power_coeff}
      \captionof{figure} {
        \label{fig:estimated_power_coeff}
        The estimated power coefficients for each mel filterbank channels using
        \eqref{eq:alpha_estimation}.
      }
    \end{center}
    \vspace{-7mm}
  \end{Figure}
%
%
%
\subsection{Histogram-based maximization of distribution uniformity}
\begin{itemize}
\item The distribution is uniformized using the empirical CDF (Refer to the paper for mathematical details):
\begin{align}
  \mathbf{Y} = \sigma_{np}(\mathbf{X}) 
     = F_u^{-1} \left(\hat{F}_x(\mathbf{X})\right)
      = \hat{F}_x(\mathbf{X}).
  \label{eq:histogram_mud_simple}
\end{align}
\end{itemize}
%
%
\begin{minipage}[]{0.54\linewidth}
  \begin{Figure}
  \begin{center}
    {
      \includegraphics[width=170mm]{mel_filterbank_energy_histogram}
      \label{fig:mel_filterbank_energy_histogram} \\
      (a)
    }   
    \\  
    {
      \includegraphics[width=170mm]{power_function_based_mud_output_histogram}
      \label{fig:power_function_based_mud_output_histogram} \\
      (b)
    }   
    \\  
    {
      \includegraphics[width=170mm]{histogram_based_mud_output_histogram}
      \label{fig:histogram_based_mud_output_histogram} \\
      (c)
    }   
    \\
    \captionof{figure} {
      The Probability Density Functions for the third filterbank channel
       of (a): the mel filterbank energy $p[m, l], l = 3$ 
       , (b): the power-function based MUD 
      output of this mel filterbank energy in \eqref{eq:power_mud}, (c) 
      the histogram based-MUD in \eqref{eq:histogram_mud_simple}. 
      \label{fig:pdf_mud}
    }   
   \vspace{-7mm}
  \end{center}
\end{Figure}
\end{minipage}
%
%
\begin{minipage}[]{0.45\linewidth}
  \begin{Figure}
  \begin{center}
  \resizebox{150mm}{!}{\input{../figures/e2e_asr_block_diagram_poster.tex}}
  \end{center}
  \captionof{figure}{ 
  The structure of the entire end-to-end speech recognition system.}
  \end{Figure}
\end{minipage}

%
%\begin{center}
%  \includegraphics[width=0.2\linewidth]{power_function_based_mud_output_histogram}

%\captionof{figure}{
%  \label{fig:bilinear_transform}
%      The relation between the input frequency and the
%      warped frequency using the bilinear transformation
%      \eqref{eq:bilinear_transformation}.}
%
%\end{center}

%\item \textbf{IFFT and OverLap Addition (OLA)} Re-synthesize the waveform.
%\end{enumerate}

%\begin{figure}%[!tbp]
%\end{figure}

\section{End-to-End Speech recognition}
 We used the RETURNN speech recognition system \cite{a_zeyer_interspeech_2018_00} with various modifications:
    \begin{itemize}
      \item{MoCha\cite{k_kim_asru_2019_00} and the modified beam search decoder.}
      \item{Gradient clipping, modified learning-late warm-up, and so on.}
    \end{itemize}

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	MATERIALS AND METHODS
%----------------------------------------------------------------------------------------

\section{Experimental Results}

\begin{itemize}
\item \textbf{LibriSpeech experiments} (960-hr training and evaluation on LibriSpeech {\tt test-clean} and {\tt test-other} ) 
\end{itemize}

\begin{minipage}[b]{1.0\linewidth}
  %\begin{table*}[!htbp]
  \renewcommand{\arraystretch}{1.3}
  \centering
      \captionof{table}{\label{tbl:result}
      Word Error Rates (WERs) obtained with MFCC, Power Mel filterbank
      coefficients,  power function-based MUD processing, and
      histogram-based MUD Processing on the LibriSpeech corpus
      \cite{v_panayotov_icassp_2015_00}. 
      }
      %\vspace{-3mm}
      \begin{tabular}{| c | c  || c | c | c | c |}
        \hline
        \multicolumn{2}{| l ||}{Neural Network Structure}
                               & \specialcell{MFCC}
                               & \specialcell{$ ( \cdot ) ^ {\frac{1}{15}}$}
                               & \specialcell{Power Function- \\ Based MUD}
                               & \specialcell{Histogram- \\ Based MUD}   \\
        \hline \hline
        \multirow{3}{*}{\specialcell{1024 cell \\ ULSTM \\ MoCha}}
                & test-clean  &   7.09  \% & \textcolor{blue}{\textbf{ 7.04 }} \%   &   7.10 \%  &  7.13 \%  \\
                & test-other  &  20.60  \% & 19.76 \%
                &  \textcolor{blue}{\textbf{19.64}} \%  & 20.03 \%  \\
                & average     &  13.85  \% & 13.40 \%
                &  \textcolor{blue}{\textbf{13.37}} \%  & 13.58 \%  \\
        \hline \hline

            \multirow{3}{*}{\specialcell{1536 cell \\ BLSTM \\ Full-Attention}}
                & test-clean  &   4.06  \% & \textcolor{blue}{\textbf{3.94}} \%   &   4.02 \%  &  4.11 \%  \\
                & test-other  &  13.97  \% & 13.56 \%
                & \textcolor{blue}{\textbf{13.34}} \%  & 14.10 \%  \\
                & average     &   9.02  \% &  8.75 \%
                & \textcolor{blue}{\textbf{8.68}} \%  &  9.11 \%  \\
        \hline
     \end{tabular}
     \vspace{-4mm}
\end{minipage}
\begin{itemize}
\item The Power-Function Based MUD shows slightly better result than the power-law nonlinearity with the power coefficient of $\frac{1}{15}$ \cite{
C_Kim_IEEETran_2016_1}.
\item We hypothesize that the reason why histogram-based MUD does slightly worse than the power-
function based MUD is that it somewhat obscured the energy boundary between speech vs non-speech.
\end{itemize}
 %----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

%\nocite{*} % Print all references regardless of whether they were cited in the poster or not
%\bibliographystyle{plain} % Plain referencing style
%\bibliography{sample} % Use the example bibliography file sample.bib

%

\color{DarkSlateGray} % DarkSlateGray color for the rest of the content
%----------------------------------------------------------------------------------------
\small
\bibliographystyle{IEEEtran}
\bibliography{../../common_bib_file/common_bib_file}
\end{multicols}
\end{document}